### 2016-05-31 Rédaction/Manuscript.tex
Départ

\documentclass[preprint,3p, authoryear]{elsarticle}

\journal{Journal of International Economics}

\begin{document}

\begin{frontmatter}

\title{Comment: Inferring Trade Costs from Trade Booms and Trade Busts}

\author[rvt]{Guillaume Corlay}
\author[rvt]{Stephane Dupraz\corref{cor1} \fnref{fn1}}
\author[rvt]{Claire Labonne}
\author[rvt]{Anne Muller}
\author[rvt]{Guillaume Daudin}
\cortext[cor1]{Corresponding author}
\ead{guillaume.daudin@dauphine.fr} \fntext[fn1]{Present
affiliation: Columbia University. Personal address : 535 W.
111th street, apt 64, 10025 New York NY. Phone number:
646-320-9026.}

\address[rvt]{ENSAE, French National School of Statistics and Economic Administration, 3, avenue Pierre Larousse, 92245 Malakoff Cedex, FRANCE}

\title{Comment: Inferring Trade Costs from Trade Booms and Trade Busts}

$\date{March 2, 2012}

\begin{abstract}
\cite{JMN2011} offer an alternative to price gaps to quantify
trade costs. Their method consists in deducing international
trade costs from trade flows. They use it to calculate the
evolution of these costs during the first and second waves of
globalization (1870-1913 and 1950-2000) and the interwar period
(1921-1939). They reach the conclusion that the reduction in
trade costs was the main driving force of the growth of trade
during the first globalization, whereas the trade boom was led
by economic expansion during the second globalization. We argue
that this result is entirely driven by the ad hoc aggregation
method of trade costs the authors use. We show that what their
measure captures is instead a difference in the size of the
countries concerned by the phenomenon of globalization. More
generally, we cast doubts on the possibility to reach
conclusions of such nature with a method that infers trade
costs from trade flows, and then uses these costs to explain
trade flows, if not to rephrase the information already
contained in an openness ratio.
\end{abstract}

\begin{keyword}
Trade costs, globalization, gravity model, aggregation,
structure effect.
\end{keyword}

\end{frontmatter}

\section{Introduction}

\cite{JMN2011} suggest\footnote{The method is also developed in
\cite{JMN2008}, \cite{JMN2010}, \cite{NOVY}.} a method for
inferring international trade costs from trade flows. They
define trade costs as all barriers to trade, notably transport
and transaction costs. Using the general equilibrium model of
\cite{AW2003}, the authors calculate those costs and their
evolution during the first and second waves of globalization
(1870-1913 and 1950-2000) as well as the interwar period
(1921-1939) thanks to the impressive set of data they collected
on trade flows, GDP and exports between 27
countries\footnote{Argentina, Australia, Austria, Belgium,
Brazil, Canada, Denmark, France, Germany, Greece, India,
Indonesia, Italy, Japan, Mexico, the Netherlands, New Zealand,
Norway, the Philippines, Portugal, Spain, Sri Lanka, Sweden,
Switzerland, the United Kingdom, the United States, and
Uruguay. The data contain 130 country pairs.}. They provide a
decomposition of the growth of trade caused by the reduction in
trade costs and economic expansion. Their main result concerns
a difference of nature between the two globalizations:

\begin{quote}
\emph{``Our results assign an overarching role for our trade
cost measure in the nineteenth century and the interwar trade
bust. In contrast, when explaining the post-World War II trade
boom, we identify a more muted role for the trade cost
measure.'' (p. 196)}.
\end{quote}
\begin{quote}
\emph{``For the pre-World War I period, we find that declines
in the trade cost measure explain roughly 60\% of the growth in
global trade. [...] Conversely, we find that only 31\% of the
present-day global trade boom can be explained by the decline
in the trade cost measure. [...] The contribution of the two
trade booms suggests that major technological breakthrough in
the nineteenth century such as the steamship, the telegraph,
and refrigeration may have been relatively more important than
technological innovations in the second half of the twentieth
century such as containerization and enhanced handling
facilities.'' (p. 186)}.
\end{quote}

We argue that this result is actually driven by the ad hoc
method of aggregation. When relying instead on a
three-countries version of \cite{AW2003}'s model for
aggregating bilateral trade costs - where the three countries
are the country of interest, its aggregate trade partners
present in the data, and the rest of the world - the alleged
difference of natures between the two globalizations fades
away. We show that the authors' main results is not robust to
the aggregation technique because the one they use also
captures structure effects. The inferred difference between the
two globalization concerns the size of the countries comes from
the phenomenon of globalization, not trade costs. Moreover, we
cast doubt on the possibility to reach conclusions of this
nature through an approach that relies on the study of trade
flows only. Deducing trade costs from trade flows, and using
trade costs to explain trade flows is essentially a circular
reasoning. Therefore, Jacks et al.'s approach cannot be an
alternative to traditional investigations of impediments to
trade, such as commodity price gaps.

This paper starts by presenting Jacks et al.'s approach to the
measure of trade costs, and insists on its relevance at the
bilateral level to control for multilateral trade barriers in
gravity regressions. It then highlights how paradoxical the
result on a difference of nature between the two globalizations
is since it cannot be deducted from a comparison of the
evolution of openness ratios. Section \ref{2} shows that the
conclusion is only driven by the authors' ad hoc aggregation
method. We propose a microfounded way to aggregate trade costs
and the puzzle fades away. Section \ref{3} explores the reasons
why Jacks et al.'s aggregation technique ends up providing
different results for the two globalizations. We argue that
what Jacks et al. misleadingly attribute to unequal trade costs
decreases between the two globalizations is instead a
difference in the distribution of trade growth over trading
partners.

\section{\label{1} Deducing trade costs from trade flows, explaining trade flows with trade costs}

Although it is consistent with many models of international
trade, Jacks, Meissner and Novy's work is primarily based on
the general equilibrium model framework of \cite{AW2003}. $n$
countries, each represented by a maximizing consumer, exchange
goods over one single period. In this Armington world,
production is not modelized\footnote{In the working paper
version of \cite{JMN2010}, the authors provide a version of the
model with production. The key equation is identical to the one
of the demand-side model.} and each country is initially
endowed with a differentiated representative good. Trade occurs
because of consumers' taste for diversity. The preferences of
all countries are assumed to be identical and modelized by a
Constant Elasticity of Substitution (CES) utility function.

\cite{AW2003} use this model to microfound gravity equations
and solve \cite{MAC}'s border puzzle by highlighting that
bilateral trade does not depend on bilateral trade barriers per
se, but bilateral trade barriers relative to impediments to
trade with all other trading partners. \cite{AW2003} show that
equilibrium imposes the following relation:

\begin{eqnarray}
x_{ij}=\frac{y_i y_j}{y^W}\left(\frac{t_{ij}}{P_i P_j}\right)^{1-\sigma}
\end{eqnarray}

where $x_{ij}$ are real exports from $i$ to $j$, $y_i$ is real
output of country $i$, $y^W$ is the world real output, $\sigma$
is the elasticity of substitution, $P_j$ is the price index in
country $j$ and can be interpreted as multilateral trade
barriers, and $t_{ij}$ is the trade costs factor between $i$
and $j$. Trade costs factors are assumed to be symmetric, i.e.
$t_{ij}$ = $t_{ji}$.

Jacks et al. first depart from \cite{AW2003} by eliminating the
multilateral resistance variable from the gravity equation.
Following \cite{NOVY}, bilateral trade barriers are not
considered relative to multilateral trade barriers modelized by
the price index, but to domestic trade costs. In this case,
trade flows are no longer compared to outputs, but to internal
trade $x_{ii}$\footnote{Due to data limitations, the authors
use the relation $x_{ii}=GDP_i-EXPORTS_i$ to get internal
trade. We follow them. Concerns about the fact that GDP is
measured in added-value and exportations as gross value are
addressed in appendix B of \cite{JMN2011}.}. The equation above
then becomes:

\begin{eqnarray}
\left( \frac{x_{ii} x_{jj}}{x_{ij} x_{ji}}\right)^{\frac{1}{2(\sigma-1)}} =\left(\frac{t_{ij} t_{ji}}{t_{ii} t_{jj}} \right)^{\frac{1}{2}} = 1+\tau_{ij} \label{keyeq}
\end{eqnarray}

Where the last equality defines $\tau_{ij}$, the geometric mean
of trade costs from country $i$ to $j$ and $j$ to $i$
\emph{relative to intra-national trade costs in countries $i$
and $j$} (no assumption of symmetry of bilateral trade costs is
imposed). It is the trade cost measure used by the authors.\\

This simple step offers a significant breakthrough in the
microfoundations of gravity equations. Comparing bilateral
trade flows to intranational trade allows to integrate
\cite{AW2003}'s caveat against omitted variable bias while
getting rid of multilateral trade barriers. This is important
when multilateral trade barriers cannot be estimated, e.g. when
we have only data on trade of one country. \cite{JMN2011} take
advantage of this feature in section 5 of their article in
regressing their measure of trade costs between two countries
on a set of proxies for trade costs, such as the distance
between trade partners, tariffs, or the volatility of the
exchange rate.

However, Jacks, Meissner and Novy suggest in their article that
\cite{AW2003}'s model can be used to move away from the gravity
regression approach. Instead of explaining trade flows by
observable proxies for costs, they quantify the impact of all
impediments to trade: \emph{``We, therefore, infer trade costs
from trade flows. This approach allows us to capture the
combined magnitude of tariffs, transport costs, and all other
macroeconomic frictions that impede international market
integration but which are inherently difficult to observe. We
emphasize that this approach of inferring trade costs from
readily available trade data holds clear advantages for applied
research: the constraints on enumerating «--» let alone,
collecting data on «--» every individual trade cost element
even over short periods of time makes a direct accounting
approach impossible.''} (p.131).

We cast doubts on the idea that a method that use trade flows
alone can teach us much about their determinants. When
departing from \cite{AW2003}'s multilateral resistances, the
arbitrage condition Jacks et al. base their work on is an
equality between a Marginal Rate of Substitution (MRS) and a
price ratio for a CES utility function, or more precisely the
product of two such equalities, one from the program of each
trade partner. Appendix \ref{A1} derives equation (\ref{keyeq})
this way without passing through multilateral resistance. The
inference of trade costs from trade flows is tantamount to
using the relation between quantities and prices given by a
demand curve derived from fully specified exogenous
preferences.

This is of course no reason in itself to disregard the method.
But when the measure is used to answer \emph{``the central
question of what drive trade booms and busts''}(p. 186) as the
authors do in section 6 of their article, the reasoning looks
circular: first calculate trade costs from trade flows, then
explain trade flows with the inferred trade costs.

Formally, the authors take the logarithm of the key-equation of
their article (\ref{keyeq}) to decompose the product of
bilateral trade flows between $i$ and $j$ in four terms:

\begin{eqnarray}
ln (x_{ij}x_{ji})= 2 ln(y_i + y_j)+ ln (s_i s_j) + 2(1-\sigma)ln(1+\tau_{ij})+ \left(\frac{x_{ii} x_{jj}}{y_i y_j}\right)
\end{eqnarray}

Where $s_i=y_i/(y_i+y_j)$. By calculating the GDP-weighted
average of the evolution of these terms over the first and
second globalizations, the authors attribute the evolution of
trade flows to four components: output growth, increasing
income similarity, changes in trade costs and a fourth term
that can be interpreted as a trade diversion effect.

Since the trade costs measure is assimilated to trade flows
relative to domestic activity, they should intuitively only
reformulate an information already contained in any measure of
trade flows relative to economic size, like an openness ratio
defined for instance as the ratio of exportations to GDP.
However the main conclusion in Jacks et al.'s article, namely
that the first globalization was more driven by the decrease in
trade costs than the second one seems to suggest that the
circularity intuition is wrong. Decomposing the level of
exportations of a country as the product of its GDP and its
openness ratio yields the results presented\footnote{We follow
the authors in displaying the changes in the log of variables
in percentage. However, one should keep in mind that given the
size of the changes, they should not be taken as percentage
variations.} in table \ref{OR}. The increase in exportations of
all countries in the table is dominantly due to the growth of
GDP for the large majority of the 27 countries present in the
data. On average the decomposition attributes 74\% (183/246) of
the growth in trade to the increase in GDP in the second
globalization and 60\% in the first globalization. To the
contrary, Jacks et al. find that their measure of trade costs
explains about 60\% (107/178) of the growth in international
trade during the first globalization, the increase in GDP
explaining about\footnote{Contributions of growth in income
similarity and of change in multilateral factors are negative.}
45\%.

\begin{table}
\centering
\begin{tabular}{|r|r|r|r|r|r|r|}
\hline
 & \multicolumn{3}{|c|}{First Globalization 1870-1913} & \multicolumn{3}{|c|}{Second Globalization 1950-2000}\\
\hline
 & Exports & GDP & Openness Ratio & Exports & GDP & Openness Ratio\\
\hline
Argentina & 292 & 251 & 40 & 116 & 132 & -16\\
\hline
Austria & 89 & 102 & -13 & 379 & 185 & 194\\
\hline
Australia & 174 & 152 & 22 & 114 & 191 & -77\\
\hline
Belgium & 195 & 86 & 109 & 251 & 150 & 101\\
\hline
Brazil & 174 & 101 & 73 & 184 & 239 & -55\\
\hline
Canada & 233 & 170 & 64 & 258 & 192 & 66\\
\hline
Denmark & 200 & 113 & 87 & 230 & 142 & 88\\
\hline
France & 117 & 70 & 48 & 260 & 172 & 87\\
\hline
Germany & 168 & 119 & 48 & 391 & 175 & 216\\
\hline
Greece & 140 & 99 & 42 & 227 & 218 & 9\\
\hline
India & 147 & 41 & 105 & 161 & 216 & -54\\
\hline
Indonesia & 220 & 87 & 133 & 263 & 232 & 31\\
\hline
Italy & 151 & 83 & 69 & 335 & 188 & 147\\
\hline
Japan & 337 & 104 & 233 & 448 & 281 & 167\\
\hline
Mexico & 189 & 143 & 46 & 296 & 238 & 59\\
\hline
Netherlands & 230 & 92 & 138 & 298 & 173 & 125\\
\hline
New Zealand & 184 & 186 & -2 & 307 & 133 & 174\\
\hline
Norway & . & . & . & 70 & 181 & -111\\
\hline
Philippines & 218 & 92 & 126 & 237 & 213 & 24\\
\hline
Portugal & 74 & 57 & 17 & 319 & 208 & 111\\
\hline
Spain & 171 & 76 & 96 & 496 & 230 & 266\\
\hline
Sri Lanka & 172 & 92 & 81 & 39 & 201 & -161\\
\hline
Sweden & 150 & 92 & 58 & 240 & 134 & 106\\
\hline
Switzerland & 105 & 108 & -3 & 250 & 132 & 118\\
\hline
UK & 125 & 81 & 44 & 196 & 122 & 74\\
\hline
Uruguay & 261 & 165 & 96 & 48 & 94 & -46\\
\hline
USA & 208 & 166 & 42 & 241 & 170 & 72\\
\hline
 &  &  &  &  &  & \\
\hline
Average & \textbf{178} & \textbf{107} & \textbf{71} & \textbf{246} & \textbf{183} & \textbf{63}\\
\hline
\end{tabular}
\caption{Decomposition of the growth of exportations between
GDP growth and Openness ratio growth, logarithms (percentages).
Figures for Norway are not given because the dissolution of the
union between Norway and Sweden (1905) make them
meaningless.}\label{OR}
\end{table}

We argue in the remaining of this comment that the first
intuition is right and that the difference between the two
globalizations Jack et al. present is only driven by their ad
hoc aggregation technique of trade costs.

\section{\label{2} Ad hoc aggregation of trade costs}

Jacks et al.'s conclusion on a difference of nature between the
two globalizations is based on an aggregate trade costs
measure. It aims at providing a summary statistic of the
evolution of trade costs across all dyads in the sample, as
well as for different regions of the world. To move from
bilateral costs $\tau_{ij}$ to an aggregate measure of trade
costs, the authors use an arithmetic mean over dyads, weighted
by the sum of the GDP of the two trade partners.

If the attractiveness of the model the authors use at the
bilateral level comes partly from its microfoundation, there is
no justification for their aggregation method. We argue that
all the results they reach that are not a reformulation of the
evolution of the openness ratio during the two
globalizations come from this aggregation technique.\\

\cite{AW2003}'s model can be used to calculate the aggregate
trade costs a single country is facing in its trade with all
its trading partners. Indeed, using a two-countries model where
all trade partners of country $i$ are treated as one single
country provides an aggregate measure of trade costs that does
not rely on any ad hoc aggregation method but instead on
microfoundations. Since not all trade flows data are available,
it is also possible to use a three-countries model, with
country $i$, its trading partners present in the data, and the
rest of the world.

Let us note $\upsilon_{i}$ the corresponding trade cost faced
by country $i$ with all its trading partners in the data. Its
expression is given by Jacks et al.'s key equation
(\ref{keyeq}) applied to a three-countries model:

\begin{eqnarray}
1+\upsilon_{i}=\left(\frac{x_{ii} x_{pp}}{x_{ip}
x_{pi}}\right)^{\frac{1}{2(\sigma-1)}} \label{keyeq3model}
\end{eqnarray}

where $x_{ip}$ and $x_{pi}$ are respectively real exports and
imports of $i$ from and to its trade partners in the data, and
$x_{pp}$ is the volume of trade within and between the trading
partners present in the data. This `domestic trade' variable
now
includes cross-border trade.\\

Using this measure of trade costs, we provide the same
decomposition in four terms of the increase in trade flows
during the two globalizations as in the authors' article for
the countries with the most trade partners available in the
data (France (24), the UK (25), the USA (23)). We also provide
the unweighted and end-of-period-GPD-weighted averages for all
countries in the sample. Results are displayed in tables
\ref{glob1} and \ref{glob2} along with the results with the
authors' method. They include unweighted and
end-of-period-GPD-weighted averages over country dyads of
bilateral trade costs, such as provided in \cite{JMN2011}. We
also display for France, the UK and the USA the trade cost
measure averaged over trading partners faced by one country, as
the results are presented (with a smaller data set) in
\cite{JMN2008, JMN2010}. We also provide both averages of this
measure over all countries in the data set.\\

\begin{table}
\centering
\begin{tabular}{|l|c|c|c|c|c|}
\hline
\vbox{\hbox{\strut 1870-1913}\hbox{\strut }\hbox{\strut}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut output}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut income similarity}} &\vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut trade cost measure}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut multilateral factors}} & \vbox{\hbox{\strut Average growth}\hbox{\strut of international}\hbox{\strut trade}}\\
\hline
&\multicolumn{5}{|c|}{JMN 2011, unweighted}\\
\hline
& 195 & 3 & 251 & -14 & 435\\
\hline
    &\multicolumn{5}{|c|}{JMN 2011, GDP-weighted}\\
\hline
& \textbf{225} & \textbf{-11} & \textbf{290} & \textbf{-18} & \textbf{486}\\
\hline
 &  &  &  &  & \\
\hline
&\multicolumn{5}{|c|}{JMN by country, unweighted}\\
\hline
France & 154 & 24 & 168 & -17 & 329\\
\hline
UK & 168 & 18 & 122 & -18 & 290\\
\hline
USA & 312 & -51 & 322 & -15 & 568\\
\hline
Average & 222 & -7 & 263 & -17 & 461\\
\hline
 &   &  &  &  & \\
\hline
&\multicolumn{5}{|c|}{JMN by country, GDP-weighted} \\
\hline
France & 187 & -1 & 151 & -12 & 325\\
\hline
UK & 194 & 3 & 102 & -13 & 286\\
\hline
USA & 286 & -29 & 244 & -11 & 490\\
\hline
Average & 228 & -9 & 295 & -13 & 501\\
\hline
 &  &  &  &  & \\
\hline
&\multicolumn{5}{|c|}{3 countries model}\\
\hline
France & 205 & -30 & 84 & -8 & 251\\
\hline
UK & 204 & -18 & 92 & -9 & 269\\
\hline
USA & 201 & 48 & 123 & -8 & 364\\
\hline
\vbox{\hbox{\strut Unweighted}\hbox{\strut average}}  & 206 & 3 & 128 & -12 & 325\\
\hline
\vbox{\hbox{\strut GDP-weighted}\hbox{\strut average}}  & \textbf{203} & \textbf{8} & \textbf{163} & \textbf{-9} & \textbf{365}\\
\hline
\end{tabular}
\caption{Decomposition of the growth in international trade
(logarithms) with ad hoc averages and a microfounded
aggregation method. First wave of globalization,
1870-1913. \emph{ \emph{JMN 2011} refers to the averaging over dyads,
\emph{JMN by country} by country refers to the averaging over
trading partners for one country, \emph{3 countries model}
refers to the aggregation method we offer.}}\label{glob1}
\end{table}

\begin{table}
\centering
\begin{tabular}{|l|c|c|c|c|c|}
\hline
\vbox{\hbox{\strut 1950-2000}\hbox{\strut }\hbox{\strut}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut output}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut income similarity}} &\vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut trade cost measure}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut multilateral factors}} & \vbox{\hbox{\strut Average growth}\hbox{\strut of international}\hbox{\strut trade}}\\
\hline
&\multicolumn{5}{|c|}{JMN 2011, unweighted}\\
\hline
 & 353 & 8 & 148 & -25 & 484\\
\hline
&\multicolumn{5}{|c|}{JMN 2011, GDP-weighted}\\
\hline
 & \textbf{350} & \textbf{3} & \textbf{137} & \textbf{-17} & \textbf{473}\\
\hline
 &  &  &  &  & \\
\hline
&\multicolumn{5}{|c|}{JMN by country, unweighted}\\
\hline
France & 355 & -1 & 201 & -29 & 526\\
\hline
UK & 280 & 29 & 54 & -25 & 338\\
\hline
USA & 343 & 11 & 112 & -21 & 445\\
\hline
Average & 349 & 9 & 124 & -23 & 459\\
\hline
 &  &  &  &  & \\
\hline
&\multicolumn{5}{|c|}{JMN by country, GDP-weighted}\\
\hline
France & 365 & 2 & 271 & -21 & 617\\
\hline
UK & 319 & 1 & 202 & -20 & 502\\
\hline
USA & 349 & 27 & 165 & -16 & 525\\
\hline
Average & 361 & 21 & 190 & -19 & 553\\
\hline
 &  &  &  &  & \\
\hline
&\multicolumn{5}{|c|}{3 countries model} \\
\hline
France & 371 & -12 & 259 & -16 & 602\\
\hline
UK & 372 & -59 & 122 & -15 & 420\\
\hline
USA & 372 & -7 & 195 & -12 & 548\\
\hline
\vbox{\hbox{\strut Unweighted}\hbox{\strut average}}  & 343 & 11 & 116 & -18 & 452\\
\hline
\vbox{\hbox{\strut GDP-weighted}\hbox{\strut average}}  & \textbf{361} & \textbf{15} & \textbf{183} & \textbf{-14} & \textbf{545}\\
\hline
\end{tabular}
\caption{Decomposition of the growth in international trade
(logarithms) with ad hoc averages and a microfounded
aggregation method. Second wave of globalization, 1950-2000.}\label{glob2}
\end{table}

These decompositions aim at deciding which of the two driving
forces - the decrease in trade costs or the income growth -
explains the increase in trade flows. The results highlight how
dependent on the aggregation technique Jacks et al.'s
decomposition is. It is actually the only cause of their
conclusion on a difference of nature between the two
globalizations. Focusing on the comparison between the results
in \cite{JMN2011} and the decomposition with a three-countries
model (in bold), the two globalizations show very similar
patterns with the latter microfounded technique. Growth in
output is the main driving force behind both waves of growth in
international trade, contributing to about 56\% (203/365) in
the first wave and 66\% (361/545) in the second one.

We selected the (GDP-weighted) average of our measure in order
to allow a clear comparison with Jacks et al.'s results. There
is of course no rational for such a summary statistic, but the
results for France, the UK and the USA assure that the main
conclusion of this exercice does not depend on averaging over
countries.

Moreover, these results strongly parallel the decomposition of
the growth of international trade between the evolutions of GDP
and openness ratio in table \ref{OR}. Insofar as a measure of
trade costs is defined residually as everything that explains
differences between domestic and international trade flows, it
is bound to reword the information contained in an openness
ratio. More ambitious conclusions are likely to be spuriously
driven by any manipulation of the data that is not backed by
the model.

\section{\label{3} Sensitivity of the trade cost measure to
structure effects}

Conclusions can go beyond this reformulation if they concern
the dispersion of the evolution of trade flows across trade
partners. We argue that if Jacks et al.'s aggregate measure of
trade costs does more than reflecting the information given by
the openness ratio, it is because aggregating trade costs by a
(unweighted or weighted) mean is sensitive to structure
effects.

To explain this idea, and for clarity purposes, let us consider
simplified expressions of aggregate trade costs by using only
the ratio of domestic trade of $i$ to imports from $j$ to $i$
$x_{ii}/x_{ji}$ for the calculation of bilateral trade costs,
when the exact expression uses the geometric average of this
ratio with the corresponding one for $j$.

\begin{eqnarray}
1+\tau_{ij}= \left(\frac{x_{ii}}{x_{ji}}\right)^{\frac{1}{\sigma-1}}\label{tausimplified}
\end{eqnarray}

This relation does not of course correspond to the arbitrage
condition of country $i$, which would involve supply prices,
but our aim here is to highlight the mechanism behind Jacks et
al.'s aggregation. Noting $\overline{a_j}$ the arithmetic mean
of $a_j$ over $j$, $\overline{a_j}=1/n \sum_{j=1}^{n}a_j$, the
simplified expressions of the unweighted average of $\tau_{ij}$
of trade partner $j$, and of the measure derives from the
three-countries model (\ref{keyeq3model}) are:

\begin{eqnarray}
1+\tau_i&=& x_{ii}^{\frac{1}{\sigma-1}} \left(\overline{{x_{ji}^{\frac{1}{1-\sigma}}}}\right) \label{tau}\\
1+\upsilon_i&=&x_{ii}^{\frac{1}{\sigma-1}} (n \times \overline{x_{ji}})^{\frac{1}{1-\sigma}}\label{upsilon}
\end{eqnarray}

Except for the factor $n$, which is irrelevant since we are
concerned with the evolution of the trade costs index, the two
expressions (\ref{tau}) and (\ref{upsilon}) differ only by the
mean they use. On the one hand, $\upsilon_i$ uses the
arithmetic mean of importations $x_{ji}$. On the other hand,
$\tau_i$, because it uses an arithmetic mean over $\tau_{ij}$,
uses a mean power $1/(1-\sigma)$ of the importations,
$\left(\overline{x_{ji}^{\frac{1}{1-\sigma}}}\right)^{1-\sigma}$.
It is established in appendix \ref{A2} that the curvature
properties of the function $x \rightarrow
x^{\frac{1}{1-\sigma}}$, $\sigma>1$ tend to draw the growth of
$\tau_i$ toward the values of the growth of bilateral trade
costs incurred with small\footnote{The `size' of a country is
used here to refer to the importance of the volume of its trade
with $i$ among trade partners of country $i$, regardless of its
population and GDP.} trading partners. To the contrary,
$\upsilon_i$ puts more weight on big trade partners, consistent
with the higher share of imports they represent. If the
arithmetic mean is independent of the distribution of the
importations over trade partners, the mean power $1/(1-\sigma)$
is extremely sensitive to structure effects.

We considered here the unweighted average of the bilateral
trade costs $\tau_{ij}$. The authors use a end-of-period-GDP
weighted average. As tables \ref{glob1} and \ref{glob2} show,
weighting by the GDP of trade partners, besides not being
theoretically justified, does not provide an accurate
correction of the bias in the measure. One reason for this
failure is that the relationship between GDP and the importance
of trade flows is not systematic.

To sum up, the (weighted or unweigthed) average of the
bilateral trade costs $\tau_{ij}$ captures two elements: trade
costs and the distribution of these costs relative to the size
of the trading partners. Through time, the more concentrated on
the small countries the growth of trade is, the more
overestimated the
decline of trade costs measured by $\tau_i$ becomes.\\

This formal discussion suggests that the strongest decline in
trade costs during the first globalization when aggregating
according to Jacks et al.'s method corresponds to a growth of
trade more concentrated on the initially small trade partners
during the first globalization.

This fits the data. For both globalizations, figure
\ref{structure} plots the growth of trade flows, as measured by
$\Delta ln (\sqrt{x_{ij} x_{ji}})$, as a function of the
initial value of trade, as measured by the logarithm of the
geometric average of bilateral average $ln(\sqrt{x_{ij}
x_{ji}})$ for all dyads in the sample. It confirms that the
growth of international trade in the first globalization is
concentrated on pairs of countries that initially traded little
together. This is what misled Jacks, Meissner and Novy into
stating a difference of nature between the two globalizations.

\begin{figure}
\centering
\includegraphics{structure.eps}
\caption{Growth of trade flows (logarithms) depending on their initial value for the first (1870-1913) and second (1950-2000) globalizations (logarithms).}\label{structure}
\end{figure}

\section{\label{5} Conclusion}

Jacks, Meissner and Novy's method for inferring trade costs
from trade flows alone simply reformulates the evolution of the
openness ratio when it is used to calculate aggregate trade
costs. This is because it only relates the two through an
equality between MRS and price ratio. That appears more clearly
when comparing the ad hoc aggregation of bilateral trade costs
to an aggregated measure of trade costs based on
\cite{AW2003}'s model. Still, if their method fails to offer a
full alternative to traditional investigations of trade costs,
such as commodity price gaps, it improves on the
microfoundation of gravity equations by substantially
simplifying the way of correcting the omitted variable bias.

This characteristic of the proposed measure of trade costs is
partly due to the comprehensively inclusive definition of trade
costs used by the authors. In such a model, costs are anything
that causes consumptions from different countries' products not
to be equal. To be sure, it is essential to highlight that
trade costs cannot be reduced to tariffs or transportation
costs, and to insist on the need for a quantification of all
the impediments to international trade. But the concept of
trade costs loses part of its interest if there are no causal
alternatives to explain trade patterns. Such a definition is
therefore bound to always reword the information given by trade
flows relative to output, such as contained in an openness
ratio.

\section*{Acknowledgments}

We are grateful to Celine Antonin (OFCE) and Guillaume Daudin
(EQUIPPE (Lille I) \& OFCE) for their constant supply of
helpful comments. They, of course, bear no responsibility for
remaining errors.

\appendix

\section{\label{A1}Appendix: Deducing the measure of trade costs from an equality between a MRS and a price ratio}

We derive in this appendix the key equation in \cite{JMN2011}.
We do not start from \cite{AW2003}'s results as we want to
highlight it is an equality between MRS and price ratio, or
more precisely the product of two such equalities, one from the
program of each trade partner.

Noting $C_{ki}$ the consumption by country $i$ of good from
region $k$, $\sigma$ the elasticity of substitution, and
$\beta_k$ a positive distribution parameter, preferences of
countries $i$ and $j$'s representative consumers are given by
their respective utility functions:

\begin{eqnarray*}
U_i=\left(\sum_k\beta_{k}^{\frac{1-\sigma}{\sigma}}C_{ki}^{\frac{\sigma-1}{\sigma}}\right)^{\frac{\sigma}{\sigma-1}}\\
U_j=\left(\sum_k\beta_{k}^{\frac{1-\sigma}{\sigma}}C_{kj}^{\frac{\sigma-1}{\sigma}}\right)^{\frac{\sigma}{\sigma-1}}
\end{eqnarray*}

Trade costs imply that prices are specific to the country of
consumption. We note $p_k$ the supply price of the producer in
country $k$ net of trade costs, and $p_{ki}$ the price of
region $k$ goods for region $i$ consumers. We define
$t_{ki}=p_{ki}/p_{k}$ the trade costs factor between $k$ and
$i$. Jacks et al.'s variable of trade costs between $k$ and
$i$, $\tau_{ki}$ is then defined as the trade costs factor
between $k$ and $i$ relative to the domestic trade costs factor
$t_{ii}$: $\tau_{ki}=t_{ki}/t_{ii}-1$. In all that follows, we
use the same notations $\tau_{ki}$ for the geometric average of
$\tau_{ki}$ and $\tau_{ik}$. Symmetry of bilateral trade costs
does not need to be assumed.

Country $i$ seeks to maximize $U_i$ under the constraint
$\sum_k p_{ki} C_{ki}=y_i$, where $y_i$ is the output of
country $i$. The first order conditions equate the Marginal
Rates of Substitution (MRS) and the price ratio. In particular,
for the goods produced by $i$ and $j$:

\begin{eqnarray*}
MRS_{ji}=\frac{\frac{\partial U_i}{\partial C_{ji}}}{\frac{\partial
U_i}{\partial C_{ii}}}=
\left(\frac{\beta_j}{\beta_i}\right)^{\frac{1-\sigma}{\sigma}}\left(\frac{C_{ji}}{C_{ii}}\right)^{-\frac{1}{\sigma}}=\frac{p_{ji}}{p_{ii}}
\end{eqnarray*}

Or equivalently with nominal values of trade flows,
$x_{ji}=p_{ji}C_{ii}$ and $x_{ii}=p_{ii}C_{ii}$ :

\begin{eqnarray*}
\left(\frac{\beta_j}{\beta_i}\right)^{\frac{1-\sigma}{\sigma}}\left(\frac{x_{ji}}{x_{ii}}\right)^{-\frac{1}{\sigma}}=\left(\frac{p_{ji}}{p_{ii}}\right)^{\frac{\sigma-1}{\sigma}}
\end{eqnarray*}

We take the power $\sigma$ of this equation. Country $k$'s
first-order optimality condition is given by a permutation of
the indexes. We hence have the system of equations:

\begin{eqnarray*}
\left(\frac{\beta_j}{\beta_i}\right)^{1-\sigma}\left(\frac{x_{ii}}{x_{ji}}\right)=\left(\frac{p_{ji}}{p_{ii}}\right)^{\sigma-1}\\
\left(\frac{\beta_i}{\beta_j}\right)^{1-\sigma}\left(\frac{x_{jj}}{x_{ij}}\right)=\left(\frac{p_{ij}}{p_{jj}}\right)^{\sigma-1}
\end{eqnarray*}

We can get rid of the $\beta_j$ coefficients by taking the
side-by-side product of those two equations:

\begin{eqnarray*}
\left(\frac{x_{ii} x_{jj}}{x_{ij} x_{ji}}\right)=\left(\frac{p_{ij}
p_{ji}}{p_{ii} p_{jj}}\right)^{\sigma-1}
\end{eqnarray*}

The main insight from this derivation is that when introducing
the trade costs variable $\tau_{ij}$, the exporter's supply
prices disappear so that the trade cost factor can be expressed
as a function of trade flows only. We get the equation on which
Jacks, Meissner and Novy base their analysis:

\begin{eqnarray*}
x_{ij} x_{ji} =(x_{ii} x_{jj}) (1+\tau_{ij})^{2(1-\sigma)}
\end{eqnarray*}

\section{\label{A2}Appendix: Properties of means and structure effects}

Let $\phi$ be a continuous bijective function. We can define
$m_{\phi}$ the $\phi$-mean of a sample $(a_j)_{1\leq j\leq n}$
as the image by $\phi^{-1}$ of the arithmetic mean of the image
of the sample $(\phi(a_j))_{1\leq j\leq n}$. A common case is
when $\phi$ is a power function $x \rightarrow x^\alpha$. It
includes the arithmetic ($\alpha=1$), quadratic ($\alpha=2$)
and harmonic ($\alpha=-1$) means. Formally:

\begin{eqnarray*}
m_{\phi}=\phi^{-1}\left(\frac{1}{n}\sum_{j=1}^n \phi(a_j)\right)
\end{eqnarray*}

Properties of these means can be deduced from the monotonicity
and convexity properties of the function $\phi$. For instance,
if $\phi$ is convex, the convexity inequality gives:

\begin{eqnarray*}
\phi(m_{\phi})=\frac{1}{n}\sum_{j=1}^n\phi(a_j)\geq \phi\left(\frac{1}{n}\sum_{j=1}^n
a_j\right)
\end{eqnarray*}

If $\phi$, and hence $\phi^{-1}$, is also increasing, the
$\phi$-mean is superior to the arithmetic mean. Intuitively,
the convexity of $\phi$ gives, relative to the arithmetic mean,
more weight to the high values in the sample. Those results are
reversed if $\phi$ is concave and increasing or convex and
decreasing. Jacks, Meissner and Novy's aggregation method of
the $\tau_{kj}$ is tantamount to using a $\phi$-mean with
$\phi$ the power functions $\phi: x \rightarrow
x^{\frac{1}{1-\rho}}$, $\rho>1$ where the coherence with the
model imposes an arithmetic mean. Hence, since $\phi$ is
decreasing and convex, the substitution tends to overweight
small countries.\\

This is a static result. We are interested in the dynamic
behavior of trade flows. But Jacks et al.'s measure also
overweights initially small trade partners in terms of the
increase in trade flows. This can be seen by calculating the
elasticities to the importations from country $k$ ($a_k$) of
the arithmetic mean ($m$) and of Jacks et al.'s mean
($m_{\sigma}$) (keeping importations from other trade partners
$a_j$ constant). The elasticity of a generic mean $m_\sigma$ to
$a_k$ is given by
\mbox{$\varepsilon_{m_{\phi}}^k=\frac{1}{n}\frac{\phi'(a_k)}{\phi'(m_{\phi})}\frac{a_k}{m_{\phi}}$},
so that:

\begin{eqnarray*}
\varepsilon_m^k&=& \frac{1}{1+\sum\limits_{j \neq k} a_j/a_k}\\
\varepsilon_{m_{\sigma}}^k&=&\frac{1}{1+\sum\limits_{j\neq k} (a_j/a_k)^{\frac{1}{1-\sigma}}}
\end{eqnarray*}

The elasticity of the arithmetic mean $m$ to $a_k$ is
increasing in $a_k$ whereas the elasticity of $m_{\sigma}$ is
decreasing in $a_k$. A one-percent increase in the importations
from $k$ increases more the arithmetic mean of importations if
$k$ is initially an important importer, simply because $k$
represents a larger part of trade. But it is the opposite with
$m_{\sigma}$: the smaller the initial value of trade with $k$,
the bigger the impact of its growth on $m_{\sigma}$. Therefore,
Jacks et al.'s measure is biased toward the growth rates of
costs incurred with small trade partners.

\nocite{*}
\bibliographystyle{elsarticle-harv}
\bibliography{bibliography}

\end{document}

### 2016-05-31 Rédaction/Manuscript.tex
Cela compile !

@@
-\documentclass[preprint,3p, authoryear]{elsarticle}
+\documentclass{article}
+\usepackage{authblk}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage[pdftex]{graphicx}  
+%Langue utilisée
+\usepackage[english]{babel}
+\usepackage[T1]{fontenc}
+\emergencystretch=8em
+\usepackage[margin=1in]{geometry}
 
-\journal{Journal of International Economics}
+\title{Comment: Inferring Trade Costs from Trade Booms and Trade Busts}
+
+\author[1]{Guillaume Corlay}
+\author[2]{Stéphane Dupraz}
+\author[1]{Claire Labonne}
+\author[1]{Anne Muller}
+%\author[3]{Guillaume Daudin}
+\author[3]{Guillaume Daudin\footnote{Corresponding author: guillaume.daudin@dauphine.fr, }\footnote{We are grateful to Celine Antonin (OFCE) for her constant supply of helpful comments.
+She, of course, bears no responsibility for remaining errors.}}
+\affil[1] {ENSAE, French National School of Statistics and Economic Administration, 3, avenue Pierre Larousse, 92245 Malakoff, FRANCE}
+\affil[2] {Columbia University, 10025 New York NY, USA}
+\affil[3] {Université Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE
+Université Paris-Dauphine, PSL Research University, LEDa, UMR [225], DIAL, 75016 PARIS, FRANCE
+
+SciencesPo, Observatoire Français des Conjonctures Économiques (OFCE), 75007, PARIS, FRANCE}
+
+%\date{March 2, 2012}
 
@@
 
-\begin{frontmatter}
-
-\title{Comment: Inferring Trade Costs from Trade Booms and Trade Busts}
-
-\author[rvt]{Guillaume Corlay}
-\author[rvt]{Stephane Dupraz\corref{cor1} \fnref{fn1}}
-\author[rvt]{Claire Labonne}
-\author[rvt]{Anne Muller}
-\author[rvt]{Guillaume Daudin}
-\cortext[cor1]{Corresponding author}
-\ead{guillaume.daudin@dauphine.fr} \fntext[fn1]{Present
-affiliation: Columbia University. Personal address : 535 W.
-111th street, apt 64, 10025 New York NY. Phone number:
-646-320-9026.}
-
-\address[rvt]{ENSAE, French National School of Statistics and Economic Administration, 3, avenue Pierre Larousse, 92245 Malakoff Cedex, FRANCE}
-
-\title{Comment: Inferring Trade Costs from Trade Booms and Trade Busts}
-
-$\date{March 2, 2012}
+\maketitle
 
@@
 
-\begin{keyword}
-Trade costs, globalization, gravity model, aggregation,
+Keywords: Trade costs, globalization, gravity model, aggregation,
 structure effect.
-\end{keyword}
-
-\end{frontmatter}
+JEL Code:
 
@@
 readily available trade data holds clear advantages for applied
-research: the constraints on enumerating «--» let alone,
-collecting data on «--» every individual trade cost element
+research: the constraints on enumerating — let alone,
+collecting data on — every individual trade cost element
 even over short periods of time makes a direct accounting
@@
 
-\section*{Acknowledgments}
-
-We are grateful to Celine Antonin (OFCE) and Guillaume Daudin
-(EQUIPPE (Lille I) \& OFCE) for their constant supply of
-helpful comments. They, of course, bear no responsibility for
-remaining errors.
-
 \appendix
@@
 \nocite{*}
-\bibliographystyle{elsarticle-harv}
+\bibliographystyle{apalike}
 \bibliography{bibliography}

### 2016-06-03 Rédaction/Manuscript.tex
Redémarrage ?

--- 2016-05-31 Rédaction/Manuscript.tex
@@
 
-\cite{JMN2011} suggest\footnote{The method is also developed in
-\cite{JMN2008}, \cite{JMN2010}, \cite{NOVY}.} a method for
-inferring international trade costs from trade flows. They
-define trade costs as all barriers to trade, notably transport
-and transaction costs. Using the general equilibrium model of
-\cite{AW2003}, the authors calculate those costs and their
+Jacks and his coauthors offer in a number of papers \footnote{The method is developed in\cite{JMN2008}, \cite{JMN2010}, \cite{NOVY} and \cite{JMN2011}.
+We will use this latter paper as a referenc} a method for
+inferring international trade costs from trade flows. Using the general equilibrium model of
+\cite{AW2003}, they calculate trade costs (defined as as all barriers to trade, notably transport
+and transaction costs) and their
 evolution during the first and second waves of globalization
@@
 decomposition of the growth of trade caused by the reduction in
-trade costs and economic expansion. Their main result concerns
+trade costs and economic expansion. They use their computations to underline
 a difference of nature between the two globalizations:
@@
 
-We argue that this result is actually driven by the ad hoc
+We argue that this result is actually driven by an ad hoc
 method of aggregation. When relying instead on a
@@
 difference of natures between the two globalizations fades
-away. We show that the authors' main results is not robust to
-the aggregation technique because the one they use also
-captures structure effects. The inferred difference between the
-two globalization concerns the size of the countries comes from
-the phenomenon of globalization, not trade costs. Moreover, we
-cast doubt on the possibility to reach conclusions of this
-nature through an approach that relies on the study of trade
-flows only. Deducing trade costs from trade flows, and using
+away. \cite{JMN2011}'s aggregation technique captures structure effects. The inferred difference between the
+two globalization
+????? comes from the evolving size of countris rather that trade costs ????
+concerns the size of the countries comes from
+the phenomenon of globalization, not trade costs. More generally, we
+cast doubt on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies only on the study of trade
+flows. Deducing trade costs from trade flows, and then using
 trade costs to explain trade flows is essentially a circular
@@
 bilateral level to control for multilateral trade barriers in
-gravity regressions. It then highlights how paradoxical the
+gravity regressions (section \ref{1}). It then highlights how paradoxical the
 result on a difference of nature between the two globalizations
@@
 model with production. The key equation is identical to the one
-of the demand-side model.} and each country is initially
+of the model without production.} and each country is initially
 endowed with a differentiated representative good. Trade occurs
@@
 trade costs than the second one seems to suggest that the
-circularity intuition is wrong. Decomposing the level of
+circularity intuition is wrong. We decompose the level of
 exportations of a country as the product of its GDP and its

### 2016-06-03 Rédaction/Manuscript.tex
Soumis à Jérôme

@@
 $t_{ij}$ = $t_{ji}$.
+
+blouf
 

### 2016-06-03 Rédaction/Manuscript.tex
Soumis à Jérôme

@@
 \end{quote}
-\begin{quote}
-\emph{``For the pre-World War I period, we find that declines
-in the trade cost measure explain roughly 60\% of the growth in
-global trade. [...] Conversely, we find that only 31\% of the
-present-day global trade boom can be explained by the decline
-in the trade cost measure. [...] The contribution of the two
-trade booms suggests that major technological breakthrough in
-the nineteenth century such as the steamship, the telegraph,
-and refrigeration may have been relatively more important than
-technological innovations in the second half of the twentieth
-century such as containerization and enhanced handling
-facilities.'' (p. 186)}.
-\end{quote}
 
@@
 two globalization
+
 ????? comes from the evolving size of countris rather that trade costs ????
+
 concerns the size of the countries comes from
@@
 
-blouf
-
 Jacks et al. first depart from \cite{AW2003} by eliminating the
@@
 trade flows are no longer compared to outputs, but to internal
-trade $x_{ii}$\footnote{Due to data limitations, the authors
+trade $x_{ii}$.\footnote{Due to data limitations, the authors
 use the relation $x_{ii}=GDP_i-EXPORTS_i$ to get internal
@@
 measured in added-value and exportations as gross value are
-addressed in appendix B of \cite{JMN2011}.}. The equation above
+addressed in appendix B of \cite{JMN2011}.} The equation above
 then becomes:
@@
 and $j$} (no assumption of symmetry of bilateral trade costs is
-imposed). It is the trade cost measure used by the authors.\\
+imposed). It is the trade cost measure used by the authors.
 
@@
 between trade partners, tariffs, or the volatility of the
-exchange rate.
+exchange rate. This is a very useful approach.\\
 
-However, Jacks, Meissner and Novy suggest in their article that
+However, Jacks, Meissner and Novy also suggest in their article that
 \cite{AW2003}'s model can be used to move away from the gravity
@@
 
-We cast doubts on the idea that a method that use trade flows
+Our point is that a method that use trade flows
 alone can teach us much about their determinants. When
 departing from \cite{AW2003}'s multilateral resistances, the
-arbitrage condition Jacks et al. base their work on is an
+arbitrage condition Jacks et al. base their computation on is an
 equality between a Marginal Rate of Substitution (MRS) and a
@@
 This is of course no reason in itself to disregard the method.
-But when the measure is used to answer \emph{``the central
+But when it cannot be used to answer \emph{``the central
 question of what drive trade booms and busts''}(p. 186) as the
-authors do in section 6 of their article, the reasoning looks
+authors do in section 6 of their article. That is because the reasoning is
 circular: first calculate trade costs from trade flows, then
-explain trade flows with the inferred trade costs.
+explain trade flows from the inferred trade costs.
+
+//
 
@@
 Since the trade costs measure is assimilated to trade flows
-relative to domestic activity, they should intuitively only
-reformulate an information already contained in any measure of
+relative to domestic activity, they simply restate
+ an information already contained in any measure of
 trade flows relative to economic size, like an openness ratio
@@
 However the main conclusion in Jacks et al.'s article, namely
-that the first globalization was more driven by the decrease in
-trade costs than the second one seems to suggest that the
-circularity intuition is wrong. We decompose the level of
+that the first globalization was driven more by the decrease in
+trade costs than the second one requires the absence of circularity.
+Indeed, when we decompose the level of
 exportations of a country as the product of its GDP and its
-openness ratio yields the results presented\footnote{We follow
-the authors in displaying the changes in the log of variables
-in percentage. However, one should keep in mind that given the
-size of the changes, they should not be taken as percentage
-variations.} in table \ref{OR}. The increase in exportations of
-all countries in the table is dominantly due to the growth of
+openness ratio, we find results that are different from theirs (see table \ref{OR}).\footnote{We follow
+the authors in interpreting log differences as percentages. However, one should keep in mind that given the
+size of the changes, this is a very inexact approximation.} The increase in exportations is mainly explained by the growth of
 GDP for the large majority of the 27 countries present in the
-data. On average the decomposition attributes 74\% (183/246) of
+data. On average, this decomposition attributes 74\% (183/246) of
 the growth in trade to the increase in GDP in the second
-globalization and 60\% in the first globalization. To the
-contrary, Jacks et al. find that their measure of trade costs
-explains about 60\% (107/178) of the growth in international
-trade during the first globalization, the increase in GDP
-explaining about\footnote{Contributions of growth in income
+globalization and 60\% in the first globalization. This is very much in contrast to 
+Jacks et al. :\footnote{Contributions of growth in income
 similarity and of change in multilateral factors are negative.}
-45\%.
+
+\begin{quote}
+\emph{``For the pre-World War I period, we find that declines
+in the trade cost measure explain roughly 60\% of the growth in
+global trade. [...] Conversely, we find that only 31\% of the
+present-day global trade boom can be explained by the decline
+in the trade cost measure. [...] The contribution of the two
+trade booms suggests that major technological breakthrough in
+the nineteenth century such as the steamship, the telegraph,
+and refrigeration may have been relatively more important than
+technological innovations in the second half of the twentieth
+century such as containerization and enhanced handling
+facilities.'' (p. 186)}.
+\end{quote}
 
@@
 \caption{Decomposition of the growth of exportations between
-GDP growth and Openness ratio growth, logarithms (percentages).
+GDP growth and Openness ratio growth, log differences (interpreted as percentages).
 Figures for Norway are not given because the dissolution of the
-union between Norway and Sweden (1905) make them
+union between Norway and Sweden (1905) makes them
 meaningless.}\label{OR}
@@
 
-We argue in the remaining of this comment that the first
-intuition is right and that the difference between the two
-globalizations Jack et al. present is only driven by their ad
-hoc aggregation technique of trade costs.
+This contrast between both decompositions of the growth of exports is at the center of the argument by Jack et al.
+As the next section shows, it is actually driven by the ad hoc way they aggregate trade costs.
 
@@
 two globalizations is based on an aggregate trade costs
-measure. It aims at providing a summary statistic of the
+measure that provides a summary statistic of the
 evolution of trade costs across all dyads in the sample, as
@@
 
-If the attractiveness of the model the authors use at the
-bilateral level comes partly from its microfoundation, there is
-no justification for their aggregation method. We argue that
+There is no justification for this aggregation method. We argue that
 all the results they reach that are not a reformulation of the
@@
 \ref{glob1} and \ref{glob2} along with the results with the
-authors' method. They include unweighted and
+Jacks et al's method.\footnote{They include unweighted and
 end-of-period-GPD-weighted averages over country dyads of
@@
 \cite{JMN2008, JMN2010}. We also provide both averages of this
-measure over all countries in the data set.\\
+measure over all countries in the data set.}\\
 
@@
 \hline
-\vbox{\hbox{\strut 1870-1913}\hbox{\strut }\hbox{\strut}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut output}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut income similarity}} &\vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut trade cost measure}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut multilateral factors}} & \vbox{\hbox{\strut Average growth}\hbox{\strut of international}\hbox{\strut trade}}\\
+\vbox{\hbox{\strut 1870-1913}\hbox{\strut }\hbox{\strut}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut output}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut income similarity}} &\vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut trade cost measure}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut multilateral factors}} & \vbox{\hbox{\strut Average growth}\hbox{\strut of bilateral}\hbox{\strut trade flows}}\\
 \hline
@@
 aggregation method. First wave of globalization,
-1870-1913. \emph{ \emph{JMN 2011} refers to the averaging over dyads,
+1870-1913.  \emph{JMN 2011} refers to the averaging over dyads,
 \emph{JMN by country} by country refers to the averaging over
 trading partners for one country, \emph{3 countries model}
-refers to the aggregation method we offer.}}\label{glob1}
+refers to the aggregation method we offer.}\label{glob1}
 \end{table}
@@
 \hline
-\vbox{\hbox{\strut 1950-2000}\hbox{\strut }\hbox{\strut}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut output}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut income similarity}} &\vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut trade cost measure}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut multilateral factors}} & \vbox{\hbox{\strut Average growth}\hbox{\strut of international}\hbox{\strut trade}}\\
+\vbox{\hbox{\strut 1950-2000}\hbox{\strut }\hbox{\strut}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut output}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut income similarity}} &\vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut trade cost measure}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut multilateral factors}} & \vbox{\hbox{\strut Average growth}\hbox{\strut of bilateral}\hbox{\strut trade flows}}\\
 \hline
@@
 
-These decompositions aim at deciding which of the two driving
-forces - the decrease in trade costs or the income growth -
-explains the increase in trade flows. The results highlight how
-dependent on the aggregation technique Jacks et al.'s
-decomposition is. It is actually the only cause of their
-conclusion on a difference of nature between the two
-globalizations. Focusing on the comparison between the results
-in \cite{JMN2011} and the decomposition with a three-countries
-model (in bold), the two globalizations show very similar
-patterns with the latter microfounded technique. Growth in
-output is the main driving force behind both waves of growth in
-international trade, contributing to about 56\% (203/365) in
-the first wave and 66\% (361/545) in the second one.
-
-We selected the (GDP-weighted) average of our measure in order
+Tables \ref{glob1} and \ref{glob2} highlight how much the decomposition
+between the decrease in trade costs and the income growth depends on the aggregation technique.
+When we use the microfounded aggregation method,\footnote{We selected the (GDP-weighted) average of our measure in order
 to allow a clear comparison with Jacks et al.'s results. There
@@
 conclusion of this exercice does not depend on averaging over
-countries.
-
-Moreover, these results strongly parallel the decomposition of
-the growth of international trade between the evolutions of GDP
-and openness ratio in table \ref{OR}. Insofar as a measure of
+countries.} 
+growth in output is the main driving force behind both waves of growth in
+international trade, contributing to about 56\% (203/365) in
+the first wave and 66\% (361/545) in the second one. It is no coïncidence
+accident if these numbers look very much like those of table \ref{OR}: 60\% (71/107) in the first globalization and 74\% (183/246) in the second one.
+Insofar as a measure of
 trade costs is defined residually as everything that explains
@@
 is bound to reword the information contained in an openness
-ratio. More ambitious conclusions are likely to be spuriously
-driven by any manipulation of the data that is not backed by
-the model.
+ratio.
 
@@
 
-Conclusions can go beyond this reformulation if they concern
-the dispersion of the evolution of trade flows across trade
-partners. We argue that if Jacks et al.'s aggregate measure of
-trade costs does more than reflecting the information given by
-the openness ratio, it is because aggregating trade costs by a
-(unweighted or weighted) mean is sensitive to structure
-effects.
-
-To explain this idea, and for clarity purposes, let us consider
-simplified expressions of aggregate trade costs by using only
-the ratio of domestic trade of $i$ to imports from $j$ to $i$
-$x_{ii}/x_{ji}$ for the calculation of bilateral trade costs,
-when the exact expression uses the geometric average of this
-ratio with the corresponding one for $j$.
+Our explanation of why Jacks et al.'s aggregate measure of
+trade costs yields different conclusion 
+from table \ref{OR} is because it is sensitive to structure
+effects. To explain this idea, and for clarity purposes, let us consider
+a word where partners are symmetric (domestic trade is equal in $i$ and $j$ and $x_ij$=$x_ji$). From equation \ref{keyeq}, we have:
 
@@
 but our aim here is to highlight the mechanism behind Jacks et
-al.'s aggregation. Noting $\overline{a_j}$ the arithmetic mean
-of $a_j$ over $j$, $\overline{a_j}=1/n \sum_{j=1}^{n}a_j$, the
-simplified expressions of the unweighted average of $\tau_{ij}$
-of trade partner $j$, and of the measure derives from the
-three-countries model (\ref{keyeq3model}) are:
+al.'s aggregation. Note $\overline{a_j}$ the arithmetic mean
+of $a_j$ over $j$ ($\overline{a_j}=1/n \sum_{j=1}^{n}a_j$). One can
+then compare the unweighted average of $\tau_{ij}$ according to equation 
+(\ref{tausimplified}) and the measure derived from the
+three-countries model in equation (\ref{keyeq3model}):
 
@@
 
-Except for the factor $n$, which is irrelevant since we are
-concerned with the evolution of the trade costs index, the two
+Except for the factor $n$ (which is irrelevant since we are
+concerned with the evolution of the trade costs index) the two
 expressions (\ref{tau}) and (\ref{upsilon}) differ only by the
 mean they use. On the one hand, $\upsilon_i$ uses the
-arithmetic mean of importations $x_{ji}$. On the other hand,
+arithmetic mean of imports $x_{ji}$. On the other hand,
 $\tau_i$, because it uses an arithmetic mean over $\tau_{ij}$,
-uses a mean power $1/(1-\sigma)$ of the importations,
+uses a mean power $1/(1-\sigma)$ of the imports,
 $\left(\overline{x_{ji}^{\frac{1}{1-\sigma}}}\right)^{1-\sigma}$.
-It is established in appendix \ref{A2} that the curvature
+Appendix \ref{A2} establishes that the curvature
 properties of the function $x \rightarrow
@@
 $\tau_i$ toward the values of the growth of bilateral trade
-costs incurred with small\footnote{The `size' of a country is
+costs incurred with small trading partners.\footnote{The `size' of a country is
 used here to refer to the importance of the volume of its trade
 with $i$ among trade partners of country $i$, regardless of its
-population and GDP.} trading partners. To the contrary,
+population and GDP.} To the contrary,
 $\upsilon_i$ puts more weight on big trade partners, consistent
-with the higher share of imports they represent. If the
-arithmetic mean is independent of the distribution of the
-importations over trade partners, the mean power $1/(1-\sigma)$
-is extremely sensitive to structure effects.
+with intuition and their higher share of imports.
+%If the
+%arithmetic mean is independent of the distribution of the
+%importations over trade partners, the mean power $1/(1-\sigma)$
+%is extremely sensitive to structure effects.
 
-We considered here the unweighted average of the bilateral
-trade costs $\tau_{ij}$. The authors use a end-of-period-GDP
-weighted average. As tables \ref{glob1} and \ref{glob2} show,
+The authors partly correct this biais in the average of $\tau_{ij}$ by the use a end-of-period-GDP
+weighted average. However, as tables \ref{glob1} and \ref{glob2} show,
 weighting by the GDP of trade partners, besides not being
@@
 failure is that the relationship between GDP and the importance
-of trade flows is not systematic.
+of trade flows is not systematic.\\
 
-To sum up, the (weighted or unweigthed) average of the
-bilateral trade costs $\tau_{ij}$ captures two elements: trade
+The average of the bilateral trade costs $\tau_{ij}$ captures two elements: trade
 costs and the distribution of these costs relative to the size
-of the trading partners. Through time, the more concentrated on
-the small countries the growth of trade is, the more
-overestimated the
-decline of trade costs measured by $\tau_i$ becomes.\\
-
-This formal discussion suggests that the strongest decline in
-trade costs during the first globalization when aggregating
-according to Jacks et al.'s method corresponds to a growth of
-trade more concentrated on the initially small trade partners
-during the first globalization.
-
+of the trading partners. Through time, the more small countries have a relatively faster growth of trade, the more the decline of trade costs measured by $\tau_i$ becomes is overestimated.
 This fits the data. For both globalizations, figure
-\ref{structure} plots the growth of trade flows, as measured by
-$\Delta ln (\sqrt{x_{ij} x_{ji}})$, as a function of the
-initial value of trade, as measured by the logarithm of the
+\ref{structure} plots the growth of trade flows (measured by
+$\Delta ln (\sqrt{x_{ij} x_{ji}})$) as a function of the
+initial value of trade, (measured by the logarithm of the
 geometric average of bilateral average $ln(\sqrt{x_{ij}
-x_{ji}})$ for all dyads in the sample. It confirms that the
+x_{ji}})$) for all dyads in the sample. It confirms that the
 growth of international trade in the first globalization is
 concentrated on pairs of countries that initially traded little
-together. This is what misled Jacks, Meissner and Novy into
+together. This is also apparent in the contrast between the rate of growth of exports
+(table \ref{OG}) and the rate of growth of bilateral trade flows (table \ref{glob1} and \ref{glob2}.
+This is what misled Jacks, Meissner and Novy into
 stating a difference of nature between the two globalizations.
@@
 Jacks, Meissner and Novy's method for inferring trade costs
-from trade flows alone simply reformulates the evolution of the
+from trade flows simply reformulates the evolution of the
 openness ratio when it is used to calculate aggregate trade

### 2016-06-06 Paper/Manuscript.tex
Changement avec Esther

--- 2016-06-03 Rédaction/Manuscript.tex
@@
 \author[1]{Anne Muller}
-%\author[3]{Guillaume Daudin}
 \author[3]{Guillaume Daudin\footnote{Corresponding author: guillaume.daudin@dauphine.fr, }\footnote{We are grateful to Celine Antonin (OFCE) for her constant supply of helpful comments.
@@
 trade costs. Their method consists in deducing international
-trade costs from trade flows. They use it to calculate the
-evolution of these costs during the first and second waves of
-globalization (1870-1913 and 1950-2000) and the interwar period
-(1921-1939). They reach the conclusion that the reduction in
+trade costs from trade flows. They argue that the reduction in
 trade costs was the main driving force of the growth of trade
-during the first globalization, whereas the trade boom was led
-by economic expansion during the second globalization. We argue
-that this result is entirely driven by the ad hoc aggregation
-method of trade costs the authors use. We show that what their
-measure captures is instead a difference in the size of the
-countries concerned by the phenomenon of globalization. More
-generally, we cast doubts on the possibility to reach
+during the first globalization (1870-1913), whereas economic expansion was the main driving force during the second globalization (1950-2000). We argue
+that this important result is  driven by the use of an ad hoc aggregation
+method. What \cite{JMN2011} capture is the 
+difference in the relative starting trade of dyads experiencing faster trade growth in the first and second globalization.
+More generally, we cast doubts on the possibility to reach
 conclusions of such nature with a method that infers trade
 costs from trade flows, and then uses these costs to explain
-trade flows, if not to rephrase the information already
-contained in an openness ratio.
+trade flows. It can only rephrase the information already
+contained in openness ratios.
 \end{abstract}
@@
 structure effect.
-JEL Code:
+
+JEL Code: F14, N70
 
@@
 
-Jacks and his coauthors offer in a number of papers \footnote{The method is developed in\cite{JMN2008}, \cite{JMN2010}, \cite{NOVY} and \cite{JMN2011}.
-We will use this latter paper as a referenc} a method for
-inferring international trade costs from trade flows. Using the general equilibrium model of
-\cite{AW2003}, they calculate trade costs (defined as as all barriers to trade, notably transport
-and transaction costs) and their
+Jacks and his coauthors offer in a number of papers  a method for
+inferring international trade costs from trade flows.\footnote{The method is developed in\cite{JMN2008}, \cite{JMN2010}, \cite{NOVY} and \cite{JMN2011}.
+We will use this latter paper as a reference} Using the general equilibrium model of
+\cite{AW2003}, they calculate trade costs (defined as as all barriers to trade, notably transport and transaction costs) and their
 evolution during the first and second waves of globalization
@@
 
-We argue that this result is actually driven by an ad hoc
-method of aggregation. When relying instead on a
-three-countries version of \cite{AW2003}'s model for
-aggregating bilateral trade costs - where the three countries
-are the country of interest, its aggregate trade partners
-present in the data, and the rest of the world - the alleged
-difference of natures between the two globalizations fades
-away. \cite{JMN2011}'s aggregation technique captures structure effects. The inferred difference between the
-two globalization
-
-????? comes from the evolving size of countris rather that trade costs ????
-
-concerns the size of the countries comes from
-the phenomenon of globalization, not trade costs. More generally, we
+This is potentially an important result that sheds light on the globalization processes. However, that result is actually driven by an ad hoc
+method of aggregation that captures structure effects. The inferred difference between the two globalization comes from the fact that the dyads with the 
+fastest growing trade in the first wave of globalization start with very small 
+trade. This is not the case in the second wave of globalization.
+Using a three-countries version of \cite{AW2003}'s model for
+the aggregation of bilateral trade costs shows that there was no 
+difference in nature betwee the two globalizations. More generally, we
 cast doubt on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies only on the study of trade
@@
 explain trade flows from the inferred trade costs.
-
-//
 
@@
 
-The average of the bilateral trade costs $\tau_{ij}$ captures two elements: trade
-costs and the distribution of these costs relative to the size
-of the trading partners. Through time, the more small countries have a relatively faster growth of trade, the more the decline of trade costs measured by $\tau_i$ becomes is overestimated.
+The average of the bilateral trade costs $\tau_{ij}$ captures both trade
+costs and the distribution of trade costs relative to the size
+of the trading partners.
+Through time, if dyads with small starting trade experience  faster growth of trade than others, the decline of trade costs measured by $\tau_i$ will be overestimated compared to the decline of $\upsilon_i$.
 This fits the data. For both globalizations, figure
@@
 geometric average of bilateral average $ln(\sqrt{x_{ij}
-x_{ji}})$) for all dyads in the sample. It confirms that the
-growth of international trade in the first globalization is
-concentrated on pairs of countries that initially traded little
-together. This is also apparent in the contrast between the rate of growth of exports
-(table \ref{OG}) and the rate of growth of bilateral trade flows (table \ref{glob1} and \ref{glob2}.
-This is what misled Jacks, Meissner and Novy into
-stating a difference of nature between the two globalizations.
+x_{ji}})$) for all dyads in the sample. This is also confirmed in the contrast between the rate of growth of exports
+(table \ref{OR}) and the rate of growth of bilateral trade flows (table \ref{glob1} and \ref{glob2}), as export growth was relatively faster than bilateral trade growth in the first globalization compared to the second (486/178 > 473/246).
+It confirms that pairs of countris that initially traded little together experienced faster growth of trade compared to other pairs in the first than in the second globalization.
+This in an interesting result, but not the one presented by JMN.
 
@@
 This characteristic of the proposed measure of trade costs is
-partly due to the comprehensively inclusive definition of trade
+partly due to the all-inclusive definition of trade
 costs used by the authors. In such a model, costs are anything
@@
 alternatives to explain trade patterns. Such a definition is
-therefore bound to always reword the information given by trade
+therefore bound to reword the information given by trade
 flows relative to output, such as contained in an openness
@@
 
-\nocite{*}
 \bibliographystyle{apalike}

### 2016-06-06 Paper/Manuscript.tex
Petit changement

@@
 evolution of the openness ratio during the two
-globalizations come from this aggregation technique.\\
+globalizations come from this aggregation method.\\
 

### 2016-06-07 Paper/Manuscript.tex
Changement de bibliographie ?
Changement de l'affiliation de Claire
Essai

@@
 \usepackage{authblk}
+\usepackage{natbib}
 \usepackage[utf8]{inputenc}
@@
 \author[2]{Stéphane Dupraz}
-\author[1]{Claire Labonne}
+\author[4]{Claire Labonne}
 \author[1]{Anne Muller}
@@
 
-SciencesPo, Observatoire Français des Conjonctures Économiques (OFCE), 75007, PARIS, FRANCE}
+Sciences Po, Observatoire Français des Conjonctures Économiques (OFCE), 75007, PARIS, FRANCE}
+
+\affil[4] { Paris School of Economics / Université Paris 1 Panthéon Sorbonne -- ACPR - Banque de France, 75009, PARIS, FRANCE}
 
@@
 
-\bibliographystyle{apalike}
+\bibliographystyle{aer}
 \bibliography{bibliography}

### 2016-07-28 Paper/Manuscript.tex
essai

@@
 
-Jacks and his coauthors offer in a number of papers  a method for
+Jacks and his coauthors offer in several papers  a method for
 inferring international trade costs from trade flows.\footnote{The method is developed in\cite{JMN2008}, \cite{JMN2010}, \cite{NOVY} and \cite{JMN2011}.

### 2016-07-28 Paper/Manuscript.tex
Corrections texte
Mes corrections du 28 juillet 2016

@@
 \author[2]{Stéphane Dupraz}
-\author[4]{Claire Labonne}
+\author[3]{Claire Labonne}
 \author[1]{Anne Muller}
-\author[3]{Guillaume Daudin\footnote{Corresponding author: guillaume.daudin@dauphine.fr, }\footnote{We are grateful to Celine Antonin (OFCE) for her constant supply of helpful comments.
-She, of course, bears no responsibility for remaining errors.}}
+\author[4]{Céline Antonin}
+\author[5]{Guillaume Daudin\footnote{Corresponding author: guillaume.daudin@dauphine.fr}
 \affil[1] {ENSAE, French National School of Statistics and Economic Administration, 3, avenue Pierre Larousse, 92245 Malakoff, FRANCE}
 \affil[2] {Columbia University, 10025 New York NY, USA}
-\affil[3] {Université Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE
+\affil[3] { Paris School of Economics / Université Paris 1 Panthéon Sorbonne -- ACPR - Banque de France, 75009, PARIS, FRANCE}
+\affil[4] {Sciences Po, Observatoire Français des Conjonctures Économiques (OFCE), 75007, PARIS, FRANCE}
+\affil[5] {Université Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE
 Université Paris-Dauphine, PSL Research University, LEDa, UMR [225], DIAL, 75016 PARIS, FRANCE
@@
 Sciences Po, Observatoire Français des Conjonctures Économiques (OFCE), 75007, PARIS, FRANCE}
-
-\affil[4] { Paris School of Economics / Université Paris 1 Panthéon Sorbonne -- ACPR - Banque de France, 75009, PARIS, FRANCE}
 
@@
 \cite{JMN2011} offer an alternative to price gaps to quantify
-trade costs. Their method consists in deducing international
-trade costs from trade flows. They argue that the reduction in
-trade costs was the main driving force of the growth of trade
+trade costs. Implementing a method which consists in deducing international
+trade costs from trade flows, they argue that the reduction in
+trade costs was the main driving force of trade growth
 during the first globalization (1870-1913), whereas economic expansion was the main driving force during the second globalization (1950-2000). We argue
-that this important result is  driven by the use of an ad hoc aggregation
+that this important result is  driven by the use of an \textit{ad hoc} aggregation
 method. What \cite{JMN2011} capture is the 
@@
 costs from trade flows, and then uses these costs to explain
-trade flows. It can only rephrase the information already
+trade flows. We argue that it can only rephrase the information already
 contained in openness ratios.
@@
 Jacks and his coauthors offer in several papers  a method for
-inferring international trade costs from trade flows.\footnote{The method is developed in\cite{JMN2008}, \cite{JMN2010}, \cite{NOVY} and \cite{JMN2011}.
+inferring international trade costs from trade flows.\footnote{The method is developed in \cite{JMN2008}, \cite{JMN2010}, \cite{NOVY} and \cite{JMN2011}.
 We will use this latter paper as a reference} Using the general equilibrium model of
-\cite{AW2003}, they calculate trade costs (defined as as all barriers to trade, notably transport and transaction costs) and their
+\cite{AW2003}, they calculate trade costs (defined as all barriers to trade, notably transportation and transaction costs) and their
 evolution during the first and second waves of globalization
@@
 
-This is potentially an important result that sheds light on the globalization processes. However, that result is actually driven by an ad hoc
+This is potentially an important result that sheds light on the globalization processes. However, that result is actually driven by an \textit{ad hoc}
 method of aggregation that captures structure effects. The inferred difference between the two globalization comes from the fact that the dyads with the 
@@
 the aggregation of bilateral trade costs shows that there was no 
-difference in nature betwee the two globalizations. More generally, we
+difference in nature between the two globalizations. More generally, we
 cast doubt on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies only on the study of trade
@@
 trade costs to explain trade flows is essentially a circular
-reasoning. Therefore, Jacks et al.'s approach cannot be an
+reasoning. Therefore, Jacks and al.'s approach cannot be an
 alternative to traditional investigations of impediments to
@@
 
-This paper starts by presenting Jacks et al.'s approach to the
+This paper starts by presenting Jacks and al.'s approach of the
 measure of trade costs, and insists on its relevance at the
 bilateral level to control for multilateral trade barriers in
-gravity regressions (section \ref{1}). It then highlights how paradoxical the
+gravity regressions (section \ref{1}). It then highlights that the
 result on a difference of nature between the two globalizations
-is since it cannot be deducted from a comparison of the
+is paradoxical since it cannot be deduced from a comparison of the
 evolution of openness ratios. Section \ref{2} shows that the
-conclusion is only driven by the authors' ad hoc aggregation
+conclusion is only driven by the authors' \textit{ad hoc} aggregation
 method. We propose a microfounded way to aggregate trade costs
 and the puzzle fades away. Section \ref{3} explores the reasons
-why Jacks et al.'s aggregation technique ends up providing
+why Jacks and al.'s aggregation technique ends up providing
 different results for the two globalizations. We argue that
-what Jacks et al. misleadingly attribute to unequal trade costs
+what Jacks and al. misleadingly attribute to unequal trade costs
 decreases between the two globalizations is instead a
@@
 and solve \cite{MAC}'s border puzzle by highlighting that
-bilateral trade does not depend on bilateral trade barriers per
-se, but bilateral trade barriers relative to impediments to
-trade with all other trading partners. \cite{AW2003} show that
+bilateral trade does not depend on bilateral trade barriers \textit{per
+se}, but bilateral trade barriers relative to impediments to
+trade with all other trading partners. \cite{AW2003} show that the
 equilibrium imposes the following relation:
@@
 output of country $i$, $y^W$ is the world real output, $\sigma$
-is the elasticity of substitution, $P_j$ is the price index in
-country $j$ and can be interpreted as multilateral trade
+is the elasticity of substitution, $P_i$ is the price index in
+country $i$ and can be interpreted as multilateral trade
 barriers, and $t_{ij}$ is the trade costs factor between $i$
@@
 
-Jacks et al. first depart from \cite{AW2003} by eliminating the
-multilateral resistance variable from the gravity equation.
+Jacks and al. first depart from \cite{AW2003} by eliminating the
+multilateral resistance variable ($P_i P_j$) from the gravity equation.
 Following \cite{NOVY}, bilateral trade barriers are not
@@
 trade flows are no longer compared to outputs, but to internal
-trade $x_{ii}$.\footnote{Due to data limitations, the authors
+trade $x_{ii}$\footnote{Due to data limitations, the authors
 use the relation $x_{ii}=GDP_i-EXPORTS_i$ to get internal
@@
 measured in added-value and exportations as gross value are
-addressed in appendix B of \cite{JMN2011}.} The equation above
-then becomes:
+addressed in appendix B of \cite{JMN2011}.}. The equation above
+becomes:
 
@@
 
-Where the last equality defines $\tau_{ij}$, the geometric mean
+The last equality defines $\tau_{ij}$, the geometric mean
 of trade costs from country $i$ to $j$ and $j$ to $i$
@@
 when multilateral trade barriers cannot be estimated, e.g. when
-we have only data on trade of one country. \cite{JMN2011} take
-advantage of this feature in section 5 of their article in
+we only have data on trade for one country. \cite{JMN2011} take
+advantage of this feature in the fifth section of their article in
 regressing their measure of trade costs between two countries
@@
 
-Our point is that a method that use trade flows
+Our point is that a method that uses trade flows
 alone can teach us much about their determinants. When
 departing from \cite{AW2003}'s multilateral resistances, the
-arbitrage condition Jacks et al. base their computation on is an
+arbitrage condition Jacks and al. base their computation on is an
 equality between a Marginal Rate of Substitution (MRS) and a
@@
 This is of course no reason in itself to disregard the method.
-But when it cannot be used to answer \emph{``the central
-question of what drive trade booms and busts''}(p. 186) as the
-authors do in section 6 of their article. That is because the reasoning is
+Yet it cannot be used to answer \emph{``the central
+question of what drives trade booms and busts''}(p. 186) as the
+authors do in section 6 of their article, because the reasoning is
 circular: first calculate trade costs from trade flows, then
@@
 
-Formally, the authors take the logarithm of the key-equation of
+Formally, the authors take the logarithm of the key equation of
 their article (\ref{keyeq}) to decompose the product of
@@
 defined for instance as the ratio of exportations to GDP.
-However the main conclusion in Jacks et al.'s article, namely
+However the main conclusion in Jacks and al.'s article, namely
 that the first globalization was driven more by the decrease in
-trade costs than the second one requires the absence of circularity.
+trade costs than the second one, requires the absence of circularity.
 Indeed, when we decompose the level of
@@
 the authors in interpreting log differences as percentages. However, one should keep in mind that given the
-size of the changes, this is a very inexact approximation.} The increase in exportations is mainly explained by the growth of
-GDP for the large majority of the 27 countries present in the
+size of the changes, this is a very inexact approximation.} The increase in exportations is mainly explained by GDP growth for the large majority of the 27 countries in the
 data. On average, this decomposition attributes 74\% (183/246) of
@@
 globalization and 60\% in the first globalization. This is very much in contrast to 
-Jacks et al. :\footnote{Contributions of growth in income
-similarity and of change in multilateral factors are negative.}
+Jacks and al.\footnote{Contributions of growth in income
+similarity and of change in multilateral factors are negative.} :
 
@@
 This contrast between both decompositions of the growth of exports is at the center of the argument by Jack et al.
-As the next section shows, it is actually driven by the ad hoc way they aggregate trade costs.
+As the next section shows, it is actually driven by the \textit{ad hoc} way they aggregate trade costs.
 
-\section{\label{2} Ad hoc aggregation of trade costs}
+\section{\label{2} \textit{Ad hoc} aggregation of trade costs}
 
-Jacks et al.'s conclusion on a difference of nature between the
+Jacks and al.'s conclusion on a difference of nature between the
 two globalizations is based on an aggregate trade costs
@@
 \cite{AW2003}'s model can be used to calculate the aggregate
-trade costs a single country is facing in its trade with all
+trade costs a single country faces in its trade with all
 its trading partners. Indeed, using a two-countries model where
@@
 country provides an aggregate measure of trade costs that does
-not rely on any ad hoc aggregation method but instead on
-microfoundations. Since not all trade flows data are available,
+not rely on any \textit{ad hoc} aggregation method but instead on
+microfoundations. Since all trade flows data are not available,
 it is also possible to use a three-countries model, with
@@
 by country $i$ with all its trading partners in the data. Its
-expression is given by Jacks et al.'s key equation
+expression is given by Jacks and al.'s key equation
 (\ref{keyeq}) applied to a three-countries model:
@@
 data (France (24), the UK (25), the USA (23)). We also provide
-the unweighted and end-of-period-GPD-weighted averages for all
+the unweighted and end-of-period-GDP-weighted averages for all
 countries in the sample. Results are displayed in tables
 \ref{glob1} and \ref{glob2} along with the results with the
-Jacks et al's method.\footnote{They include unweighted and
+Jacks and al's method.\footnote{They include unweighted and
 end-of-period-GPD-weighted averages over country dyads of
@@
 \caption{Decomposition of the growth in international trade
-(logarithms) with ad hoc averages and a microfounded
+(logarithms) with \textit{ad hoc} averages and a microfounded
 aggregation method. First wave of globalization,
@@
 \caption{Decomposition of the growth in international trade
-(logarithms) with ad hoc averages and a microfounded
+(logarithms) with \textit{ad hoc} averages and a microfounded
 aggregation method. Second wave of globalization, 1950-2000.}\label{glob2}
@@
 When we use the microfounded aggregation method,\footnote{We selected the (GDP-weighted) average of our measure in order
-to allow a clear comparison with Jacks et al.'s results. There
+to allow a clear comparison with Jacks and al.'s results. There
 is of course no rational for such a summary statistic, but the
@@
 international trade, contributing to about 56\% (203/365) in
-the first wave and 66\% (361/545) in the second one. It is no coïncidence
-accident if these numbers look very much like those of table \ref{OR}: 60\% (71/107) in the first globalization and 74\% (183/246) in the second one.
+the first wave and 66\% (361/545) in the second one. It is no coincidence
+if these results are very similar to the ones in table \ref{OR}: 60\% (71/107) in the first globalization and 74\% (183/246) in the second one.
 Insofar as a measure of
@@
 
-Our explanation of why Jacks et al.'s aggregate measure of
+Our explanation of why Jacks and al.'s aggregate measure of
 trade costs yields different conclusion 
@@
 x^{\frac{1}{1-\sigma}}$, $\sigma>1$ tend to draw the growth of
-$\tau_i$ toward the values of the growth of bilateral trade
+$\tau_i$ towards the values of the growth of bilateral trade
 costs incurred with small trading partners.\footnote{The `size' of a country is
@@
 with $i$ among trade partners of country $i$, regardless of its
-population and GDP.} To the contrary,
+population and GDP.} On the contrary,
 $\upsilon_i$ puts more weight on big trade partners, consistent
@@
 $\Delta ln (\sqrt{x_{ij} x_{ji}})$) as a function of the
-initial value of trade, (measured by the logarithm of the
+initial value of trade (measured by the logarithm of the
 geometric average of bilateral average $ln(\sqrt{x_{ij}
@@
 (table \ref{OR}) and the rate of growth of bilateral trade flows (table \ref{glob1} and \ref{glob2}), as export growth was relatively faster than bilateral trade growth in the first globalization compared to the second (486/178 > 473/246).
-It confirms that pairs of countris that initially traded little together experienced faster growth of trade compared to other pairs in the first than in the second globalization.
+It confirms that the pairs of countries that initially traded little together experienced faster growth of trade compared to other pairs in the first than in the second globalization.
 This in an interesting result, but not the one presented by JMN.
@@
 costs. This is because it only relates the two through an
-equality between MRS and price ratio. That appears more clearly
-when comparing the ad hoc aggregation of bilateral trade costs
+equality between MRS and price ratio. It appears more clearly
+when comparing the \textit{ad hoc} aggregation of bilateral trade costs
 to an aggregated measure of trade costs based on
@@
 full alternative to traditional investigations of trade costs,
-such as commodity price gaps, it improves on the
+such as commodity price gaps, it provides an improvement in the
 microfoundation of gravity equations by substantially
@@
 
-Noting $C_{ki}$ the consumption by country $i$ of good from
+Let us note $C_{ki}$ the consumption by country $i$ of good from
 region $k$, $\sigma$ the elasticity of substitution, and
@@
 $t_{ki}=p_{ki}/p_{k}$ the trade costs factor between $k$ and
-$i$. Jacks et al.'s variable of trade costs between $k$ and
+$i$. Jacks and al.'s variable of trade costs between $k$ and
 $i$, $\tau_{ki}$ is then defined as the trade costs factor
@@
 This is a static result. We are interested in the dynamic
-behavior of trade flows. But Jacks et al.'s measure also
+behavior of trade flows. But Jacks and al.'s measure also
 overweights initially small trade partners in terms of the
@@
 elasticities to the importations from country $k$ ($a_k$) of
-the arithmetic mean ($m$) and of Jacks et al.'s mean
+the arithmetic mean ($m$) and of Jacks and al.'s mean
 ($m_{\sigma}$) (keeping importations from other trade partners
@@
 the bigger the impact of its growth on $m_{\sigma}$. Therefore,
-Jacks et al.'s measure is biased toward the growth rates of
+Jacks and al.'s measure is biased toward the growth rates of
 costs incurred with small trade partners.

### 2016-07-28 Paper/Manuscript.tex
jacks et / jacks and

@@
 trade costs to explain trade flows is essentially a circular
-reasoning. Therefore, Jacks and al.'s approach cannot be an
+reasoning. Therefore, Jacks et al.'s approach cannot be an
 alternative to traditional investigations of impediments to
@@
 
-This paper starts by presenting Jacks and al.'s approach of the
+This paper starts by presenting Jacks et al.'s approach of the
 measure of trade costs, and insists on its relevance at the
@@
 and the puzzle fades away. Section \ref{3} explores the reasons
-why Jacks and al.'s aggregation technique ends up providing
+why Jacks et al.'s aggregation technique ends up providing
 different results for the two globalizations. We argue that
-what Jacks and al. misleadingly attribute to unequal trade costs
+what Jacks et al. misleadingly attribute to unequal trade costs
 decreases between the two globalizations is instead a
@@
 
-Jacks and al. first depart from \cite{AW2003} by eliminating the
+Jacks et al. first depart from \cite{AW2003} by eliminating the
 multilateral resistance variable ($P_i P_j$) from the gravity equation.
@@
 departing from \cite{AW2003}'s multilateral resistances, the
-arbitrage condition Jacks and al. base their computation on is an
+arbitrage condition Jacks et al. base their computation on is an
 equality between a Marginal Rate of Substitution (MRS) and a
@@
 defined for instance as the ratio of exportations to GDP.
-However the main conclusion in Jacks and al.'s article, namely
+However the main conclusion in Jacks et al.'s article, namely
 that the first globalization was driven more by the decrease in
@@
 globalization and 60\% in the first globalization. This is very much in contrast to 
-Jacks and al.\footnote{Contributions of growth in income
+Jacks et al.\footnote{Contributions of growth in income
 similarity and of change in multilateral factors are negative.} :
@@
 
-Jacks and al.'s conclusion on a difference of nature between the
+Jacks et al.'s conclusion on a difference of nature between the
 two globalizations is based on an aggregate trade costs
@@
 by country $i$ with all its trading partners in the data. Its
-expression is given by Jacks and al.'s key equation
+expression is given by Jacks et al.'s key equation
 (\ref{keyeq}) applied to a three-countries model:
@@
 \ref{glob1} and \ref{glob2} along with the results with the
-Jacks and al's method.\footnote{They include unweighted and
+Jacks et al.'s method.\footnote{They include unweighted and
 end-of-period-GPD-weighted averages over country dyads of
@@
 When we use the microfounded aggregation method,\footnote{We selected the (GDP-weighted) average of our measure in order
-to allow a clear comparison with Jacks and al.'s results. There
+to allow a clear comparison with Jacks et al.'s results. There
 is of course no rational for such a summary statistic, but the
@@
 
-Our explanation of why Jacks and al.'s aggregate measure of
+Our explanation of why Jacks et al.'s aggregate measure of
 trade costs yields different conclusion 
@@
 $t_{ki}=p_{ki}/p_{k}$ the trade costs factor between $k$ and
-$i$. Jacks and al.'s variable of trade costs between $k$ and
+$i$. Jacks et al.'s variable of trade costs between $k$ and
 $i$, $\tau_{ki}$ is then defined as the trade costs factor
@@
 This is a static result. We are interested in the dynamic
-behavior of trade flows. But Jacks and al.'s measure also
+behavior of trade flows. But Jacks et al.'s measure also
 overweights initially small trade partners in terms of the
@@
 elasticities to the importations from country $k$ ($a_k$) of
-the arithmetic mean ($m$) and of Jacks and al.'s mean
+the arithmetic mean ($m$) and of Jacks et al.'s mean
 ($m_{\sigma}$) (keeping importations from other trade partners
@@
 the bigger the impact of its growth on $m_{\sigma}$. Therefore,
-Jacks and al.'s measure is biased toward the growth rates of
+Jacks et al.'s measure is biased toward the growth rates of
 costs incurred with small trade partners.

### 2016-07-29 Paper/Manuscript.tex
Correction pour compilation
+ passage de "et al." en italique

@@
 \author[4]{Céline Antonin}
-\author[5]{Guillaume Daudin\footnote{Corresponding author: guillaume.daudin@dauphine.fr}
+\author[5]{Guillaume Daudin\footnote{Corresponding author: guillaume.daudin@dauphine.fr}}
 \affil[1] {ENSAE, French National School of Statistics and Economic Administration, 3, avenue Pierre Larousse, 92245 Malakoff, FRANCE}
@@
 Université Paris-Dauphine, PSL Research University, LEDa, UMR [225], DIAL, 75016 PARIS, FRANCE
-
 Sciences Po, Observatoire Français des Conjonctures Économiques (OFCE), 75007, PARIS, FRANCE}
@@
 trade costs to explain trade flows is essentially a circular
-reasoning. Therefore, Jacks et al.'s approach cannot be an
+reasoning. Therefore, Jacks \textit{et al.}'s approach cannot be an
 alternative to traditional investigations of impediments to
@@
 
-This paper starts by presenting Jacks et al.'s approach of the
+This paper starts by presenting Jacks \textit{et al.}'s approach of the
 measure of trade costs, and insists on its relevance at the
@@
 and the puzzle fades away. Section \ref{3} explores the reasons
-why Jacks et al.'s aggregation technique ends up providing
+why Jacks \textit{et al.}'s aggregation technique ends up providing
 different results for the two globalizations. We argue that
-what Jacks et al. misleadingly attribute to unequal trade costs
+what Jacks \textit{et al.} misleadingly attribute to unequal trade costs
 decreases between the two globalizations is instead a
@@
 
-Jacks et al. first depart from \cite{AW2003} by eliminating the
+Jacks \textit{et al.} first depart from \cite{AW2003} by eliminating the
 multilateral resistance variable ($P_i P_j$) from the gravity equation.
@@
 departing from \cite{AW2003}'s multilateral resistances, the
-arbitrage condition Jacks et al. base their computation on is an
+arbitrage condition Jacks \textit{et al.} base their computation on is an
 equality between a Marginal Rate of Substitution (MRS) and a
@@
 defined for instance as the ratio of exportations to GDP.
-However the main conclusion in Jacks et al.'s article, namely
+However the main conclusion in Jacks \textit{et al.}'s article, namely
 that the first globalization was driven more by the decrease in
@@
 globalization and 60\% in the first globalization. This is very much in contrast to 
-Jacks et al.\footnote{Contributions of growth in income
+Jacks \textit{et al.}\footnote{Contributions of growth in income
 similarity and of change in multilateral factors are negative.} :
@@
 
-This contrast between both decompositions of the growth of exports is at the center of the argument by Jack et al.
-As the next section shows, it is actually driven by the \textit{ad hoc} way they aggregate trade costs.
+This contrast between both decompositions of the growth of exports is at the center of the argument by Jack \textit{et al.} As the next section shows, it is actually driven by the \textit{ad hoc} way they aggregate trade costs.
 
@@
 
-Jacks et al.'s conclusion on a difference of nature between the
+Jacks \textit{et al.}'s conclusion on a difference of nature between the
 two globalizations is based on an aggregate trade costs
@@
 by country $i$ with all its trading partners in the data. Its
-expression is given by Jacks et al.'s key equation
+expression is given by Jacks \textit{et al.}'s key equation
 (\ref{keyeq}) applied to a three-countries model:
@@
 \ref{glob1} and \ref{glob2} along with the results with the
-Jacks et al.'s method.\footnote{They include unweighted and
+Jacks \textit{et al.}'s method.\footnote{They include unweighted and
 end-of-period-GPD-weighted averages over country dyads of
@@
 When we use the microfounded aggregation method,\footnote{We selected the (GDP-weighted) average of our measure in order
-to allow a clear comparison with Jacks et al.'s results. There
+to allow a clear comparison with Jacks \textit{et al.}'s results. There
 is of course no rational for such a summary statistic, but the
@@
 
-Our explanation of why Jacks et al.'s aggregate measure of
+Our explanation of why Jacks \textit{et al.}'s aggregate measure of
 trade costs yields different conclusion 
@@
 $t_{ki}=p_{ki}/p_{k}$ the trade costs factor between $k$ and
-$i$. Jacks et al.'s variable of trade costs between $k$ and
+$i$. Jacks \textit{et al.}'s variable of trade costs between $k$ and
 $i$, $\tau_{ki}$ is then defined as the trade costs factor
@@
 This is a static result. We are interested in the dynamic
-behavior of trade flows. But Jacks et al.'s measure also
+behavior of trade flows. But Jacks \textit{et al.}'s measure also
 overweights initially small trade partners in terms of the
@@
 elasticities to the importations from country $k$ ($a_k$) of
-the arithmetic mean ($m$) and of Jacks et al.'s mean
+the arithmetic mean ($m$) and of Jacks \textit{et al.}'s mean
 ($m_{\sigma}$) (keeping importations from other trade partners
@@
 the bigger the impact of its growth on $m_{\sigma}$. Therefore,
-Jacks et al.'s measure is biased toward the growth rates of
+Jacks \textit{et al.}'s measure is biased toward the growth rates of
 costs incurred with small trade partners.

### 2016-08-30 Paper/Manuscript.tex
Correction référence Novy + institutions Guillaume

@@
 \affil[5] {Université Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE
+
 Université Paris-Dauphine, PSL Research University, LEDa, UMR [225], DIAL, 75016 PARIS, FRANCE
+
 Sciences Po, Observatoire Français des Conjonctures Économiques (OFCE), 75007, PARIS, FRANCE}

### 2016-09-19 Paper/Manuscript.tex
Ré-introduction Keith and Head

@@
 multilateral resistance variable ($P_i P_j$) from the gravity equation.
-Following \cite{NOVY}, bilateral trade barriers are not
-considered relative to multilateral trade barriers modelized by
-the price index, but to domestic trade costs. In this case,
+As in \cite{NOVY}, they use the Head-Ries index (\cite{HR2001}) to express bilateral trade barriers not relatively to multilateral trade barriers modelized by
+the price index, but relatively to domestic trade costs. In this case,
 trade flows are no longer compared to outputs, but to internal

### 2017-09-13 Paper/Manuscript.tex
Points 1 et 2 de l’éditeur

@@
 
-Université Paris-Dauphine, PSL Research University, LEDa, UMR [225], DIAL, 75016 PARIS, FRANCE
-
-Sciences Po, Observatoire Français des Conjonctures Économiques (OFCE), 75007, PARIS, FRANCE}
+Université Paris-Dauphine, PSL Research University, LEDa, UMR [225], DIAL, 75016 PARIS, FRANCE}
 
@@
 \begin{abstract}
-\cite{JMN2011} offer an alternative to price gaps to quantify
-trade costs. Implementing a method which consists in deducing international
-trade costs from trade flows, they argue that the reduction in
-trade costs was the main driving force of trade growth
-during the first globalization (1870-1913), whereas economic expansion was the main driving force during the second globalization (1950-2000). We argue
-that this important result is  driven by the use of an \textit{ad hoc} aggregation
-method. What \cite{JMN2011} capture is the 
-difference in the relative starting trade of dyads experiencing faster trade growth in the first and second globalization.
-More generally, we cast doubts on the possibility to reach
-conclusions of such nature with a method that infers trade
-costs from trade flows, and then uses these costs to explain
-trade flows. We argue that it can only rephrase the information already
-contained in openness ratios.
+\cite{JMN2011} offer a method to measure trade costs that relies exclusively on bilateral exports and GDP statistics.
+They argue that the reduction in trade costs was the main driving force of trade growth during the first globalization (1870-1913), whereas economic expansion was the main driving force during the second globalization (1950-2000).
+This potentially important result is driven by the use of an \textit{ad hoc} aggregation method of bilateral trade costs to the country and global level.
+What \cite{JMN2011} capture is the difference in the relative starting trade of dyads experiencing faster trade growth in the first and second globalization.
+More generally, we cast doubts on the possibility to reach conclusions of such nature with a method that infers trade costs from trade flows, and then uses these costs to explain trade flows.
+We argue that it can only rephrase the information already contained in openness ratios.
 \end{abstract}
@@
 
-Jacks and his coauthors offer in several papers  a method for
-inferring international trade costs from trade flows.\footnote{The method is developed in \cite{JMN2008}, \cite{JMN2010}, \cite{NOVY} and \cite{JMN2011}.
+Jacks and his coauthors offer in several papers  an innovative method to measure trade costs.\footnote{The method is developed in \cite{JMN2008}, \cite{JMN2010}, \cite{NOVY} and \cite{JMN2011}.
 We will use this latter paper as a reference} Using the general equilibrium model of
@@
 
-This is potentially an important result that sheds light on the globalization processes. However, that result is actually driven by an \textit{ad hoc}
-method of aggregation that captures structure effects. The inferred difference between the two globalization comes from the fact that the dyads with the 
-fastest growing trade in the first wave of globalization start with very small 
-trade. This is not the case in the second wave of globalization.
-Using a three-countries version of \cite{AW2003}'s model for
-the aggregation of bilateral trade costs shows that there was no 
-difference in nature between the two globalizations. More generally, we
-cast doubt on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies only on the study of trade
-flows. Deducing trade costs from trade flows, and then using
-trade costs to explain trade flows is essentially a circular
-reasoning. Therefore, Jacks \textit{et al.}'s approach cannot be an
-alternative to traditional investigations of impediments to
-trade, such as commodity price gaps.
+This is potentially an important result that sheds light on the globalization processes.
+However, that result is actually driven by an \textit{ad hoc} method of aggregation that captures structure effects.
+The authors use an end-of-period GDP arithmetic average of bilateral transport costs to compute country-specific trade costs.
+In a simple symmetric three-country model, we show that this is the equivalent to weighting the information given by each dyadic trade flow by the product of end-of-period GDP.
+and the value of the trade flow to the power $1/(1-\sigma)$  (with \sigma, as discussed latter, > 1).
+In contrast, a theoretically justified method weights, more naturally, the information given by each dyadic trade flow by the the value of the trade flow. 
+Because $1/(1-\sigma)$ < 0, this means among other things that the importance of small dyadic trade flows in the computation of country-specific trade costs is too large in the authors' computations.
+Their conclusion on the difference between the two globalization comes from the fact that the dyads with the fastest growing trade in the first wave of globalization start with very small trade and this this is not the case in the second wave of globalization.
+
+In this paper, we use a three-countries version of \cite{AW2003}'s model for the aggregation of bilateral trade costs shows that there was no difference in nature between the two globalizations to explain the issue.
+More generally, we cast doubt on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies only on the study of trade flows.
+Deducing trade costs from trade flows, and then using trade costs to explain trade flows is essentially a circular reasoning.
+Therefore, Jacks \textit{et al.}'s approach cannot be an alternative to traditional investigations of impediments to trade, such as commodity price gaps.
 
@@
 
-Our point is that a method that uses trade flows
-alone can teach us much about their determinants. When
-departing from \cite{AW2003}'s multilateral resistances, the
-arbitrage condition Jacks \textit{et al.} base their computation on is an
-equality between a Marginal Rate of Substitution (MRS) and a
-price ratio for a CES utility function, or more precisely the
-product of two such equalities, one from the program of each
-trade partner. Appendix \ref{A1} derives equation (\ref{keyeq})
-this way without passing through multilateral resistance. The
-inference of trade costs from trade flows is tantamount to
-using the relation between quantities and prices given by a
-demand curve derived from fully specified exogenous
-preferences.
+Our point is that a method that uses trade flows alone can teach us so much about their determinants.
+When departing from \cite{AW2003}'s multilateral resistances, the arbitrage condition Jacks \textit{et al.} base their computation on is an equality between a Marginal Rate of Substitution (MRS) and a price ratio for a CES utility function, or more precisely the product of two such equalities, one from the program of each trade partner.
+Appendix \ref{A1} derives equation (\ref{keyeq}) this way without passing through multilateral resistance.
+The inference of trade costs from trade flows is tantamount to using the relation between quantities and prices given by a demand curve derived from fully specified exogenous preferences.
 
 This is of course no reason in itself to disregard the method.
-Yet it cannot be used to answer \emph{``the central
-question of what drives trade booms and busts''}(p. 186) as the
-authors do in section 6 of their article, because the reasoning is
-circular: first calculate trade costs from trade flows, then
-explain trade flows from the inferred trade costs.
+Yet it cannot be used to answer \emph{``the central question of what drives trade booms and busts''}(p.
+186) as the authors do in section 6 of their article, because the reasoning is circular: first calculate trade costs from trade flows, then explain trade flows from the inferred trade costs.
 
-Formally, the authors take the logarithm of the key equation of
-their article (\ref{keyeq}) to decompose the product of
-bilateral trade flows between $i$ and $j$ in four terms:
+Formally, the authors take the logarithm of the key equation of their article (\ref{keyeq}) to decompose the product of bilateral trade flows between $i$ and $j$ in four terms:
 
@@
 
-Where $s_i=y_i/(y_i+y_j)$. By calculating the GDP-weighted
-average of the evolution of these terms over the first and
-second globalizations, the authors attribute the evolution of
-trade flows to four components: output growth, increasing
-income similarity, changes in trade costs and a fourth term
-that can be interpreted as a trade diversion effect.
+Where $s_i=y_i/(y_i+y_j)$.
+The authors hence attribute the evolution of bilateral trade flows to four components: output growth, increasing income similarity, changes in trade costs and a trade diversion effect.
 
-Since the trade costs measure is assimilated to trade flows
-relative to domestic activity, they simply restate
- an information already contained in any measure of
-trade flows relative to economic size, like an openness ratio
-defined for instance as the ratio of exportations to GDP.
+Since the trade costs measure is assimilated to trade flows relative to domestic activity, they simply restate an information already contained in any measure of trade flows relative to economic size, like an openness ratio defined for instance as the ratio of exportations to GDP.
+
 However the main conclusion in Jacks \textit{et al.}'s article, namely
@@
 mean they use. On the one hand, $\upsilon_i$ uses the
-arithmetic mean of imports $x_{ji}$. On the other hand,
+arithmetic mean of trade flows $x_{ji}$. On the other hand,
 $\tau_i$, because it uses an arithmetic mean over $\tau_{ij}$,
-uses a mean power $1/(1-\sigma)$ of the imports,
+uses a mean power $1/(1-\sigma)$ of the value of trade flows,
 $\left(\overline{x_{ji}^{\frac{1}{1-\sigma}}}\right)^{1-\sigma}$.
@@
 $\upsilon_i$ puts more weight on big trade partners, consistent
-with intuition and their higher share of imports.
+with intuition and their higher share of trade flows.
 %If the

### 2017-09-13 Paper/Manuscript.tex
Bug

@@
 However, that result is actually driven by an \textit{ad hoc} method of aggregation that captures structure effects.
-The authors use an end-of-period GDP arithmetic average of bilateral transport costs to compute country-specific trade costs.
-In a simple symmetric three-country model, we show that this is the equivalent to weighting the information given by each dyadic trade flow by the product of end-of-period GDP.
-and the value of the trade flow to the power $1/(1-\sigma)$  (with \sigma, as discussed latter, > 1).
-In contrast, a theoretically justified method weights, more naturally, the information given by each dyadic trade flow by the the value of the trade flow. 
+The authors use an weighted arithmetic average of bilateral transport costs to compute country-specific trade costs.
+Putting aside the weight, in a simple symmetric three-country model, we show that this is the equivalent to using a power or generalized mean with exponent $1/(1-\sigma)$ on the value of dyadic trade flows.
+In contrast, a theoretically justified method would use a simple arithmetic mean of the value of the dyadic trade flow. 
 Because $1/(1-\sigma)$ < 0, this means among other things that the importance of small dyadic trade flows in the computation of country-specific trade costs is too large in the authors' computations.
+This is not compensated by the weight they use (end-of-period GDP).
 Their conclusion on the difference between the two globalization comes from the fact that the dyads with the fastest growing trade in the first wave of globalization start with very small trade and this this is not the case in the second wave of globalization.

### 2017-09-14 Paper/Manuscript.tex
Answering point 3

@@
 What \cite{JMN2011} capture is the difference in the relative starting trade of dyads experiencing faster trade growth in the first and second globalization.
-More generally, we cast doubts on the possibility to reach conclusions of such nature with a method that infers trade costs from trade flows, and then uses these costs to explain trade flows.
-We argue that it can only rephrase the information already contained in openness ratios.
+More generally, we cast doubts on the possibility to reach conclusions of such nature with a method that excludes a priori changes in non-trade costs determinants of openness rates and hence can only rephrases the information contained in openness ratios.
 \end{abstract}
@@
 The authors use an weighted arithmetic average of bilateral transport costs to compute country-specific trade costs.
-Putting aside the weight, in a simple symmetric three-country model, we show that this is the equivalent to using a power or generalized mean with exponent $1/(1-\sigma)$ on the value of dyadic trade flows.
+Putting aside the weight, in a simple symmetric three-country model, we show that this is the equivalent to using a power or generalized mean with exponent $1/(1-\sigma)$ on the value of dyadic trade flows ($\sigma > 1$ is elasticity of substitution).
 In contrast, a theoretically justified method would use a simple arithmetic mean of the value of the dyadic trade flow. 
-Because $1/(1-\sigma)$ < 0, this means among other things that the importance of small dyadic trade flows in the computation of country-specific trade costs is too large in the authors' computations.
+Because $1/(1-\sigma)$ < 0, the importance of small dyadic trade flows in the computation of country-specific trade costs is too large in the authors' computations.
 This is not compensated by the weight they use (end-of-period GDP).
@@
 In this paper, we use a three-countries version of \cite{AW2003}'s model for the aggregation of bilateral trade costs shows that there was no difference in nature between the two globalizations to explain the issue.
-More generally, we cast doubt on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies only on the study of trade flows.
+More generally, we cast doubt on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies only on the study of trade flows and excludes a priori other possibles causes for the evolution of openness rates, like the evolution of vertical specialization and changes in the elasticity of substitution between domestic and foreign goods.
 Deducing trade costs from trade flows, and then using trade costs to explain trade flows is essentially a circular reasoning.
@@
 trade. We follow them. Concerns about the fact that GDP is
-measured in added-value and exportations as gross value are
+measured in value-added and exportations as gross value are
 addressed in appendix B of \cite{JMN2011}.}. The equation above
@@
 
-The last equality defines $\tau_{ij}$, the geometric mean
-of trade costs from country $i$ to $j$ and $j$ to $i$
-\emph{relative to intra-national trade costs in countries $i$
-and $j$} (no assumption of symmetry of bilateral trade costs is
-imposed). It is the trade cost measure used by the authors.
+The last equality defines $\tau_{ij}$, the geometric mean of trade costs from country $i$ to $j$ and $j$ to $i$ \emph{relative to intra-national trade costs in countries $i$ and $j$} (no assumption of symmetry of bilateral trade costs is imposed).
+It is the trade cost measure used by the authors.
 
-This simple step offers a significant breakthrough in the
-microfoundations of gravity equations. Comparing bilateral
-trade flows to intranational trade allows to integrate
-\cite{AW2003}'s caveat against omitted variable bias while
-getting rid of multilateral trade barriers. This is important
-when multilateral trade barriers cannot be estimated, e.g. when
-we only have data on trade for one country. \cite{JMN2011} take
-advantage of this feature in the fifth section of their article in
-regressing their measure of trade costs between two countries
-on a set of proxies for trade costs, such as the distance
-between trade partners, tariffs, or the volatility of the
-exchange rate. This is a very useful approach.\\
-
-However, Jacks, Meissner and Novy also suggest in their article that
-\cite{AW2003}'s model can be used to move away from the gravity
-regression approach. Instead of explaining trade flows by
-observable proxies for costs, they quantify the impact of all
-impediments to trade: \emph{``We, therefore, infer trade costs
-from trade flows. This approach allows us to capture the
-combined magnitude of tariffs, transport costs, and all other
-macroeconomic frictions that impede international market
-integration but which are inherently difficult to observe. We
-emphasize that this approach of inferring trade costs from
-readily available trade data holds clear advantages for applied
-research: the constraints on enumerating — let alone,
-collecting data on — every individual trade cost element
-even over short periods of time makes a direct accounting
-approach impossible.''} (p.131).
-
-Our point is that a method that uses trade flows alone can teach us so much about their determinants.
 When departing from \cite{AW2003}'s multilateral resistances, the arbitrage condition Jacks \textit{et al.} base their computation on is an equality between a Marginal Rate of Substitution (MRS) and a price ratio for a CES utility function, or more precisely the product of two such equalities, one from the program of each trade partner.
@@
 The inference of trade costs from trade flows is tantamount to using the relation between quantities and prices given by a demand curve derived from fully specified exogenous preferences.
+This is of course no reason in itself to disregard the method.
 
-This is of course no reason in itself to disregard the method.
-Yet it cannot be used to answer \emph{``the central question of what drives trade booms and busts''}(p.
-186) as the authors do in section 6 of their article, because the reasoning is circular: first calculate trade costs from trade flows, then explain trade flows from the inferred trade costs.
+Actually, this simple step offers a significant breakthrough in the microfoundations of gravity equations. 
+Comparing bilateral trade flows to intranational trade allows to integrate \cite{AW2003}'s caveat against omitted variable bias while getting rid of multilateral trade barriers.
+This is important when multilateral trade barriers cannot be estimated, e.g.
+when we only have data on trade for one country.
+\cite{JMN2011} take advantage of this feature in the fifth section of their article in regressing their measure of trade costs between two countries on a set of proxies for trade costs, such as the distance between trade partners, tariffs, or the volatility of the exchange rate.
+This is a very useful approach.\\
+
+However, Jacks, Meissner and Novy also suggest in their article that \cite{AW2003}'s model can be used to move away from the gravity regression approach.
+Instead of explaining trade flows by observable proxies for costs, they quantify the impact of all impediments to trade:
+\emph{``We, therefore, infer trade costs from trade flows.
+This approach allows us to capture the combined magnitude of tariffs, transport costs, and all other macroeconomic frictions that impede international market integration but which are inherently difficult to observe.
+We emphasize that this approach of inferring trade costs from readily available trade data holds clear advantages for applied research: the constraints on enumerating — let alone, collecting data on — every individual trade cost element even over short periods of time makes a direct accounting approach impossible.''}(p.131).
 
@@
 Where $s_i=y_i/(y_i+y_j)$.
-The authors hence attribute the evolution of bilateral trade flows to four components: output growth, increasing income similarity, changes in trade costs and a trade diversion effect.
 
-Since the trade costs measure is assimilated to trade flows relative to domestic activity, they simply restate an information already contained in any measure of trade flows relative to economic size, like an openness ratio defined for instance as the ratio of exportations to GDP.
+The authors hence attribute the evolution of bilateral trade flows to four components: output growth, increasing total income similarity, changes in trade costs and a trade diversion effect.
 
-However the main conclusion in Jacks \textit{et al.}'s article, namely
-that the first globalization was driven more by the decrease in
-trade costs than the second one, requires the absence of circularity.
-Indeed, when we decompose the level of
-exportations of a country as the product of its GDP and its
-openness ratio, we find results that are different from theirs (see table \ref{OR}).\footnote{We follow
-the authors in interpreting log differences as percentages. However, one should keep in mind that given the
-size of the changes, this is a very inexact approximation.} The increase in exportations is mainly explained by GDP growth for the large majority of the 27 countries in the
-data. On average, this decomposition attributes 74\% (183/246) of
-the growth in trade to the increase in GDP in the second
-globalization and 60\% in the first globalization. This is very much in contrast to 
-Jacks \textit{et al.}\footnote{Contributions of growth in income
-similarity and of change in multilateral factors are negative.} :
+Examining this equation, it is clear that the method consists in determining the evolution of one unobservable variable (trade costs) based on the evolution of two groups of observable variables (trade flows and output) and a number of hypothesis on preferences and parameters ($\sigma$ and the relationship between $x_{ii}$, output and trade).
+If one assumes that preferences and parameters do not change (though these points are partly addressed  in appendix B of  \cite{JMN2011}), all changes in trade costs must restate information contained in changes in the measure of trade flows relative to economic size, like an openness ratio defined for instance as the ratio of exportations to GDP.
+
+A method that uses data on trade flows and output alone can only teach us so much about their determinants as non-trade costs determinants of trade flows (e.g.
+the elasticity of substitution and the importance of vertical specialization) are treated as parameters.
+
+Still, when we decompose the level of exportations of a country simply as the product of its GDP and its openness ratio (see table \ref{OR}) we find that the increase in exportations is mainly explained by GDP growth for the large majority of the 27 countries in the data.\footnote{We follow the authors in interpreting log differences as percentages.
+However, one should keep in mind that given the size of the changes, this is a very inexact approximation.}
+On average, this decomposition attributes 74\% (183/246) of the growth in trade to the increase in GDP in the second globalization and 60\% in the first globalization.
+This is very much in contrast to  Jacks \textit{et al.}\footnote{Contributions of growth in income similarity and of change in multilateral factors are negative.} :
 
 \begin{quote}
-\emph{``For the pre-World War I period, we find that declines
-in the trade cost measure explain roughly 60\% of the growth in
-global trade. [...] Conversely, we find that only 31\% of the
-present-day global trade boom can be explained by the decline
-in the trade cost measure. [...] The contribution of the two
-trade booms suggests that major technological breakthrough in
-the nineteenth century such as the steamship, the telegraph,
-and refrigeration may have been relatively more important than
-technological innovations in the second half of the twentieth
-century such as containerization and enhanced handling
-facilities.'' (p. 186)}.
+\emph{``For the pre-World War I period, we find that declines in the trade cost measure explain roughly 60\% of the growth in global trade.
+[...] Conversely, we find that only 31\% of the present-day global trade boom can be explained by the decline in the trade cost measure.
+[...] The contribution of the two trade booms suggests that major technological breakthrough in the nineteenth century such as the steamship, the telegraph, and refrigeration may have been relatively more important than technological innovations in the second half of the twentieth century such as containerization and enhanced handling facilities.'' (p.
+186)}.
 \end{quote}
+
+This contrast between both decompositions of the growth of exports is at the center of the argument by Jack \textit{et al.}.
+It is surprising because it is not clear what is added by their method to the examination of the openness rates.
+As the next section shows, this contrast it is actually driven by the \textit{ad hoc} way they aggregate trade costs.
 
@@
 \end{table}
-
-This contrast between both decompositions of the growth of exports is at the center of the argument by Jack \textit{et al.} As the next section shows, it is actually driven by the \textit{ad hoc} way they aggregate trade costs.
 

### 2017-09-14 Paper/Manuscript.tex
Nouvelle equation pour point 4

@@
 \begin{eqnarray}
-1+\upsilon_{i}=\left(\frac{x_{ii} x_{pp}}{x_{ip}
-x_{pi}}\right)^{\frac{1}{2(\sigma-1)}} \label{keyeq3model}
+1+\upsilon_{i}= \left(\frac{x_{ii} x_{pp}}{x_{ip}
+x_{pi}}\right)^{\frac{1}{2(\sigma-1)}}
+=\left(\frac{x_{ii} \sum_{j \neq i, k \neq i } x_{jk}}{\sum_{j \neq i } x_{ij}\sum_{j \neq i} x_{ji}}\right)^{\frac{1}{2(\sigma-1)}}   \label{keyeq3model}
 \end{eqnarray}
@@
 includes cross-border trade.\\
+
+Compare this with Jacks \textit{et al.}'s  method : 
+\begin{eqnarray}
+1+\tau_i&=& \sum_{j \neq i}  {\frac{y_i}{\sum_{j \neq i} y_i} \left(1+\tau_{ij}\right)}
+= \sum_{j \neq i}  {\frac{y_i}{\sum_{j \neq i} y_i} \left( \frac{x_{ii} x_{jj}}{x_{ij} x_{ji}}\right)^{\frac{1}{2(\sigma-1)}} }
+\end{eqnarray}
 

### 2017-09-14 Paper/Manuscript.tex
Thanks and out with 3-country method

@@
 \author[4]{Céline Antonin}
-\author[5]{Guillaume Daudin\footnote{Corresponding author: guillaume.daudin@dauphine.fr}}
+\author[5]{Guillaume Daudin\footnote{Corresponding author: guillaume.daudin@dauphine.fr. The authors thank Marc Adam, Marcelo Olarreaga and anonymous referees for their comments and questions}}
 \affil[1] {ENSAE, French National School of Statistics and Economic Administration, 3, avenue Pierre Larousse, 92245 Malakoff, FRANCE}
@@
 not rely on any \textit{ad hoc} aggregation method but instead on
-microfoundations. Since all trade flows data are not available,
-it is also possible to use a three-countries model, with
-country $i$, its trading partners present in the data, and the
-rest of the world.
+microfoundations.
 
@@
 now
-includes cross-border trade.\\
+includes cross-border trade. The problem with this method is that it is very data-intensive. It requires the knowledge of all bilateral trade flows between the countries in the data. \\
 

### 2017-09-15 Paper/Manuscript.tex
Pour tableau 1 et Our method
Comme dans le papier soumis

@@
 \usepackage[margin=1in]{geometry}
+\usepackage{booktabs}
+\usepackage{tabularx}
 
@@
 \author[4]{Céline Antonin}
-\author[5]{Guillaume Daudin\footnote{Corresponding author: guillaume.daudin@dauphine.fr. The authors thank Marc Adam, Marcelo Olarreaga and anonymous referees for their comments and questions}}
+\author[5]{Guillaume Daudin\footnote{Corresponding author: guillaume.daudin@dauphine.fr.
+The authors thank Marc Adam, Marcelo Olarreaga and anonymous referees for their comments and questions.
+Marc Adam pointed to us a slight data incoherence in Jacks \textit{et al.}'s .dta files that we have corrected in the following analysis.}}
 \affil[1] {ENSAE, French National School of Statistics and Economic Administration, 3, avenue Pierre Larousse, 92245 Malakoff, FRANCE}
@@
 What \cite{JMN2011} capture is the difference in the relative starting trade of dyads experiencing faster trade growth in the first and second globalization.
-More generally, we cast doubts on the possibility to reach conclusions of such nature with a method that excludes a priori changes in non-trade costs determinants of openness rates and hence can only rephrases the information contained in openness ratios.
+More generally, we cast doubts on the possibility to reach conclusions on aggregate costs with a method that excludes a priori changes in non-trade costs determinants of openness rates and hence can only rephrases the information contained in them.
 \end{abstract}
@@
 However, one should keep in mind that given the size of the changes, this is a very inexact approximation.}
-On average, this decomposition attributes 74\% (183/246) of the growth in trade to the increase in GDP in the second globalization and 60\% in the first globalization.
+On average, this decomposition attributes 74\% (183/246) of the growth in trade to the increase in GDP in the second globalization and 62\% in the first globalization.
 This is very much in contrast to  Jacks \textit{et al.}\footnote{Contributions of growth in income similarity and of change in multilateral factors are negative.} :
@@
 
-\begin{table}
-\centering
-\begin{tabular}{|r|r|r|r|r|r|r|}
-\hline
- & \multicolumn{3}{|c|}{First Globalization 1870-1913} & \multicolumn{3}{|c|}{Second Globalization 1950-2000}\\
-\hline
- & Exports & GDP & Openness Ratio & Exports & GDP & Openness Ratio\\
-\hline
-Argentina & 292 & 251 & 40 & 116 & 132 & -16\\
-\hline
-Austria & 89 & 102 & -13 & 379 & 185 & 194\\
-\hline
-Australia & 174 & 152 & 22 & 114 & 191 & -77\\
-\hline
-Belgium & 195 & 86 & 109 & 251 & 150 & 101\\
-\hline
-Brazil & 174 & 101 & 73 & 184 & 239 & -55\\
-\hline
-Canada & 233 & 170 & 64 & 258 & 192 & 66\\
-\hline
-Denmark & 200 & 113 & 87 & 230 & 142 & 88\\
-\hline
-France & 117 & 70 & 48 & 260 & 172 & 87\\
-\hline
-Germany & 168 & 119 & 48 & 391 & 175 & 216\\
-\hline
-Greece & 140 & 99 & 42 & 227 & 218 & 9\\
-\hline
-India & 147 & 41 & 105 & 161 & 216 & -54\\
-\hline
-Indonesia & 220 & 87 & 133 & 263 & 232 & 31\\
-\hline
-Italy & 151 & 83 & 69 & 335 & 188 & 147\\
-\hline
-Japan & 337 & 104 & 233 & 448 & 281 & 167\\
-\hline
-Mexico & 189 & 143 & 46 & 296 & 238 & 59\\
-\hline
-Netherlands & 230 & 92 & 138 & 298 & 173 & 125\\
-\hline
-New Zealand & 184 & 186 & -2 & 307 & 133 & 174\\
-\hline
-Norway & . & . & . & 70 & 181 & -111\\
-\hline
-Philippines & 218 & 92 & 126 & 237 & 213 & 24\\
-\hline
-Portugal & 74 & 57 & 17 & 319 & 208 & 111\\
-\hline
-Spain & 171 & 76 & 96 & 496 & 230 & 266\\
-\hline
-Sri Lanka & 172 & 92 & 81 & 39 & 201 & -161\\
-\hline
-Sweden & 150 & 92 & 58 & 240 & 134 & 106\\
-\hline
-Switzerland & 105 & 108 & -3 & 250 & 132 & 118\\
-\hline
-UK & 125 & 81 & 44 & 196 & 122 & 74\\
-\hline
-Uruguay & 261 & 165 & 96 & 48 & 94 & -46\\
-\hline
-USA & 208 & 166 & 42 & 241 & 170 & 72\\
-\hline
- &  &  &  &  &  & \\
-\hline
-Average & \textbf{178} & \textbf{107} & \textbf{71} & \textbf{246} & \textbf{183} & \textbf{63}\\
-\hline
-\end{tabular}
+\begin{table}[tbp] \centering
+\newcolumntype{C}{>{\centering\arraybackslash}X}
+
+\begin{tabularx}{\textwidth}{lCCCCCC}
+
+\toprule
+& \multicolumn{3}{c}{First Globalization 1870-1913} & \multicolumn{3}{c}{Second Globalization 1950-2000}\\
+{}&{Exports}&{GDP}&{Openness Ratio}&{Exports}&{GDP}&{Openness Ratio} \tabularnewline
+\midrule\addlinespace[1.5ex]
+ARG&292&251&40&116&132&-16 \tabularnewline
+AUH&89&102&-13&379&185&194 \tabularnewline
+AUS&174&152&22&114&191&-77 \tabularnewline
+BEL&195&86&109&251&150&101 \tabularnewline
+BRA&174&101&73&184&239&-55 \tabularnewline
+CAN&233&170&64&258&192&66 \tabularnewline
+DEN&200&113&87&230&142&88 \tabularnewline
+FRA&117&70&48&260&172&87 \tabularnewline
+GER&168&119&48&391&175&216 \tabularnewline
+GRE&140&99&42&227&218&9 \tabularnewline
+IND&147&41&105&161&216&-54 \tabularnewline
+INN&220&87&133&263&232&31 \tabularnewline
+ITA&151&83&69&335&188&147 \tabularnewline
+JAP&337&104&233&448&281&167 \tabularnewline
+MEX&189&143&46&296&238&59 \tabularnewline
+NET&230&92&138&298&173&125 \tabularnewline
+NEW&184&186&-2&307&133&174 \tabularnewline
+NOR&&&&70&181&-111 \tabularnewline
+PHI&218&92&126&237&213&24 \tabularnewline
+POR&74&57&17&319&208&111 \tabularnewline
+SPA&171&76&96&496&230&266 \tabularnewline
+SRI&172&92&81&39&201&-161 \tabularnewline
+SWE&150&92&58&240&134&106 \tabularnewline
+SWI&105&108&-3&250&132&118 \tabularnewline
+UK&125&81&44&196&122&74 \tabularnewline
+URU&261&165&96&48&94&-46 \tabularnewline
+USA&208&166&42&241&170&72 \tabularnewline
+\textbf{Average}&\textbf{182}&\textbf{113}&\textbf{69}&\textbf{246}&\textbf{183}&\textbf{63} \tabularnewline
+\bottomrule \addlinespace[1.5ex]
+
+\end{tabularx}
 \caption{Decomposition of the growth of exportations between
@@
 not rely on any \textit{ad hoc} aggregation method but instead on
-microfoundations.
+microfoundations. The drawback of this method is that it is very data-intensive. It requires the knowledge of all bilateral trade flows between the countries in the data.
 
@@
 now
-includes cross-border trade. The problem with this method is that it is very data-intensive. It requires the knowledge of all bilateral trade flows between the countries in the data. \\
+includes cross-border trade. \\
 
-Compare this with Jacks \textit{et al.}'s  method : 
+Compare this with Jacks \textit{et al.}'s  method to measure the mean trade cost faced by a country : 
 \begin{eqnarray}
@@
 
-Using this measure of trade costs, we provide the same
-decomposition in four terms of the increase in trade flows
-during the two globalizations as in the authors' article for
-the countries with the most trade partners available in the
-data (France (24), the UK (25), the USA (23)). We also provide
-the unweighted and end-of-period-GDP-weighted averages for all
-countries in the sample. Results are displayed in tables
-\ref{glob1} and \ref{glob2} along with the results with the
+To illustrate the practical difference between these two methods, we provide the same decomposition in four terms of the increase in trade flows during the two globalizations as in the authors' article for the countries with the most trade partners available in the data (France (24), the UK (25), the USA (23)).
+This requires studying only trade between Belgium, Canada, France, Germany, India, Italy, Spain, United Kingdom and United States of America, as this is the largest set in the data where all bilateral trade flows are available.
+We also provide the unweighted and end-of-period-GDP-weighted averages for all countries in the sample. 
+Results are displayed in tables \ref{glob1} and \ref{glob2} along with the results with the
 Jacks \textit{et al.}'s method.\footnote{They include unweighted and
 end-of-period-GPD-weighted averages over country dyads of
-bilateral trade costs, such as provided in \cite{JMN2011}. We
-also display for France, the UK and the USA the trade cost
-measure averaged over trading partners faced by one country, as
-the results are presented (with a smaller data set) in
-\cite{JMN2008, JMN2010}. We also provide both averages of this
-measure over all countries in the data set.}\\
+bilateral trade costs, such as provided in \cite{JMN2011}.
+We also display for France, the UK and the USA the trade cost measure averaged over trading partners faced by one country, as the results are presented (with a smaller data set) in
+\cite{JMN2008, JMN2010}.
+We also provide both averages of this measure over all countries in the data set.}\\
 

### 2017-09-15 Paper/Manuscript.tex
Pour les tableaux dans le texte

@@
 
-\begin{table}
-\centering
-\begin{tabular}{|l|c|c|c|c|c|}
-\hline
-\vbox{\hbox{\strut 1870-1913}\hbox{\strut }\hbox{\strut}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut output}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut income similarity}} &\vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut trade cost measure}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut multilateral factors}} & \vbox{\hbox{\strut Average growth}\hbox{\strut of bilateral}\hbox{\strut trade flows}}\\
-\hline
-&\multicolumn{5}{|c|}{JMN 2011, unweighted}\\
-\hline
-& 195 & 3 & 251 & -14 & 435\\
-\hline
-    &\multicolumn{5}{|c|}{JMN 2011, GDP-weighted}\\
-\hline
-& \textbf{225} & \textbf{-11} & \textbf{290} & \textbf{-18} & \textbf{486}\\
-\hline
- &  &  &  &  & \\
-\hline
-&\multicolumn{5}{|c|}{JMN by country, unweighted}\\
-\hline
-France & 154 & 24 & 168 & -17 & 329\\
-\hline
-UK & 168 & 18 & 122 & -18 & 290\\
-\hline
-USA & 312 & -51 & 322 & -15 & 568\\
-\hline
-Average & 222 & -7 & 263 & -17 & 461\\
-\hline
- &   &  &  &  & \\
-\hline
-&\multicolumn{5}{|c|}{JMN by country, GDP-weighted} \\
-\hline
-France & 187 & -1 & 151 & -12 & 325\\
-\hline
-UK & 194 & 3 & 102 & -13 & 286\\
-\hline
-USA & 286 & -29 & 244 & -11 & 490\\
-\hline
-Average & 228 & -9 & 295 & -13 & 501\\
-\hline
- &  &  &  &  & \\
-\hline
-&\multicolumn{5}{|c|}{3 countries model}\\
-\hline
-France & 205 & -30 & 84 & -8 & 251\\
-\hline
-UK & 204 & -18 & 92 & -9 & 269\\
-\hline
-USA & 201 & 48 & 123 & -8 & 364\\
-\hline
-\vbox{\hbox{\strut Unweighted}\hbox{\strut average}}  & 206 & 3 & 128 & -12 & 325\\
-\hline
-\vbox{\hbox{\strut GDP-weighted}\hbox{\strut average}}  & \textbf{203} & \textbf{8} & \textbf{163} & \textbf{-9} & \textbf{365}\\
-\hline
-\end{tabular}
+\begin{table}[tbp] \centering
+\newcolumntype{C}{>{\centering\arraybackslash}X}
+
+\begin{tabularx}{\textwidth}{lCCCCC}
+
+\toprule
+{}&{Contribution of growth in output}&{Contribution of growth in income similarity}&{Contribution of change in trade cost measure}&{Contribution of change in multilateral factors}&{Average growth of bilateral trade flows} \tabularnewline
+\midrule\addlinespace[1.5ex]
+&\multicolumn{5}{c}{JMN 2011} \tabularnewline
+GDP-weighted&215&-6&275&-13&470 \tabularnewline
+Unweighted&193&5&283&-15&466 \tabularnewline
+\midrule &\multicolumn{5}{c}{JMN by country, unweighted} \tabularnewline
+FRA&165&7&122&-13&281 \tabularnewline
+UK&175&7&96&-14&264 \tabularnewline
+USA&289&-33&195&-11&441 \tabularnewline
+Average&216&-7&272&-12&469 \tabularnewline
+\midrule &\multicolumn{5}{c}{JMN by country, GDP-weighted} \tabularnewline
+FRA&196&-11&127&-10&302 \tabularnewline
+UK&201&-4&69&-11&256 \tabularnewline
+USA&274&-25&192&-8&433 \tabularnewline
+Average&227&-12&254&-10&460 \tabularnewline
+\midrule &\multicolumn{5}{c}{Our method} \tabularnewline
+FRA&202&-27&77&-6&245 \tabularnewline
+UK&202&-16&67&-8&244 \tabularnewline
+USA&202&43&89&-7&327 \tabularnewline
+Unweighted average&202&-3&137&-7&329 \tabularnewline
+GDP-weighted average&202&13&126&-7&333 \tabularnewline
+\bottomrule \addlinespace[1.5ex]
+
+\end{tabularx}
 \caption{Decomposition of the growth in international trade
@@
 \emph{JMN by country} by country refers to the averaging over
-trading partners for one country, \emph{3 countries model}
+trading partners for one country, \emph{our methodl}
 refers to the aggregation method we offer.}\label{glob1}
@@
 
-\begin{table}
-\centering
-\begin{tabular}{|l|c|c|c|c|c|}
-\hline
-\vbox{\hbox{\strut 1950-2000}\hbox{\strut }\hbox{\strut}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut output}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of growth in}\hbox{\strut income similarity}} &\vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut trade cost measure}} & \vbox{\hbox{\strut Contribution}\hbox{\strut of change in}\hbox{\strut multilateral factors}} & \vbox{\hbox{\strut Average growth}\hbox{\strut of bilateral}\hbox{\strut trade flows}}\\
-\hline
-&\multicolumn{5}{|c|}{JMN 2011, unweighted}\\
-\hline
- & 353 & 8 & 148 & -25 & 484\\
-\hline
-&\multicolumn{5}{|c|}{JMN 2011, GDP-weighted}\\
-\hline
- & \textbf{350} & \textbf{3} & \textbf{137} & \textbf{-17} & \textbf{473}\\
-\hline
- &  &  &  &  & \\
-\hline
-&\multicolumn{5}{|c|}{JMN by country, unweighted}\\
-\hline
-France & 355 & -1 & 201 & -29 & 526\\
-\hline
-UK & 280 & 29 & 54 & -25 & 338\\
-\hline
-USA & 343 & 11 & 112 & -21 & 445\\
-\hline
-Average & 349 & 9 & 124 & -23 & 459\\
-\hline
- &  &  &  &  & \\
-\hline
-&\multicolumn{5}{|c|}{JMN by country, GDP-weighted}\\
-\hline
-France & 365 & 2 & 271 & -21 & 617\\
-\hline
-UK & 319 & 1 & 202 & -20 & 502\\
-\hline
-USA & 349 & 27 & 165 & -16 & 525\\
-\hline
-Average & 361 & 21 & 190 & -19 & 553\\
-\hline
- &  &  &  &  & \\
-\hline
-&\multicolumn{5}{|c|}{3 countries model} \\
-\hline
-France & 371 & -12 & 259 & -16 & 602\\
-\hline
-UK & 372 & -59 & 122 & -15 & 420\\
-\hline
-USA & 372 & -7 & 195 & -12 & 548\\
-\hline
-\vbox{\hbox{\strut Unweighted}\hbox{\strut average}}  & 343 & 11 & 116 & -18 & 452\\
-\hline
-\vbox{\hbox{\strut GDP-weighted}\hbox{\strut average}}  & \textbf{361} & \textbf{15} & \textbf{183} & \textbf{-14} & \textbf{545}\\
-\hline
-\end{tabular}
+\begin{table}[tbp] \centering
+\newcolumntype{C}{>{\centering\arraybackslash}X}
+
+\begin{tabularx}{\textwidth}{lCCCCC}
+
+\toprule
+{}&{Contribution of growth in output}&{Contribution of growth in income similarity}&{Contribution of change in trade cost measure}&{Contribution of change in multilateral factors}&{Average growth of bilateral trade flows} \tabularnewline
+\midrule\addlinespace[1.5ex]
+&\multicolumn{5}{c}{JMN 2011} \tabularnewline
+GDP-weighted&350&6&230&-28&558 \tabularnewline
+Unweighted&355&4&284&-35&608 \tabularnewline
+\midrule &\multicolumn{5}{c}{JMN by country, unweighted} \tabularnewline
+FRA&350&3&318&-30&641 \tabularnewline
+UK&297&11&227&-29&507 \tabularnewline
+USA&341&9&180&-23&507 \tabularnewline
+Average&350&6&224&-27&554 \tabularnewline
+\midrule &\multicolumn{5}{c}{JMN by country, GDP-weighted} \tabularnewline
+FRA&347&2&259&-21&587 \tabularnewline
+UK&315&-12&219&-20&502 \tabularnewline
+USA&342&10&161&-18&495 \tabularnewline
+Average&347&9&184&-21&519 \tabularnewline
+\midrule &\multicolumn{5}{c}{Our method} \tabularnewline
+FRA&347&-1&317&-16&648 \tabularnewline
+UK&347&-46&236&-15&522 \tabularnewline
+USA&347&0&192&-13&526 \tabularnewline
+Unweighted average&347&7&254&-16&592 \tabularnewline
+GDP-weighted average&347&2&203&-14&538 \tabularnewline
+\bottomrule \addlinespace[1.5ex]
+
+\end{tabularx}
 \caption{Decomposition of the growth in international trade
 (logarithms) with \textit{ad hoc} averages and a microfounded
-aggregation method. Second wave of globalization, 1950-2000.}\label{glob2}
+aggregation method. Second wave of globalization, 1950-2000.
+\emph{JMN 2011} refers to the averaging over dyads,
+\emph{JMN by country} by country refers to the averaging over
+trading partners for one country, \emph{our methodl}
+refers to the aggregation method we offer.}\label{glob2}
 \end{table}

### 2017-09-15 Paper/Manuscript.tex
Changement du texte et des tables

@@
 \toprule
-{}&{Contribution of growth in output}&{Contribution of growth in income similarity}&{Contribution of change in trade cost measure}&{Contribution of change in multilateral factors}&{Average growth of bilateral trade flows} \tabularnewline
+{1870-1913}&{Contribution of growth in output}&{Contribution of growth in income similarity}&{Contribution of change in trade cost measure}&{Contribution of change in multilateral factors}&{Average growth of bilateral trade flows} \tabularnewline
 \midrule\addlinespace[1.5ex]
 &\multicolumn{5}{c}{JMN 2011} \tabularnewline
-GDP-weighted&215&-6&275&-13&470 \tabularnewline
 Unweighted&193&5&283&-15&466 \tabularnewline
+\textbf{GDP-weighted}&\textbf{215}&\textbf{-6}&\textbf{275}&\textbf{-13}&\textbf{470} \tabularnewline
+
 \midrule &\multicolumn{5}{c}{JMN by country, unweighted} \tabularnewline
@@
 Unweighted average&202&-3&137&-7&329 \tabularnewline
-GDP-weighted average&202&13&126&-7&333 \tabularnewline
+\textbf{GDP-weighted average}&\textbf{202}&\textbf{13}&\textbf{126}&\textbf{-7}&\textbf{333} \tabularnewline
 \bottomrule \addlinespace[1.5ex]
@@
 \toprule
-{}&{Contribution of growth in output}&{Contribution of growth in income similarity}&{Contribution of change in trade cost measure}&{Contribution of change in multilateral factors}&{Average growth of bilateral trade flows} \tabularnewline
+{1950-2000}&{Contribution of growth in output}&{Contribution of growth in income similarity}&{Contribution of change in trade cost measure}&{Contribution of change in multilateral factors}&{Average growth of bilateral trade flows} \tabularnewline
 \midrule\addlinespace[1.5ex]
 &\multicolumn{5}{c}{JMN 2011} \tabularnewline
-GDP-weighted&350&6&230&-28&558 \tabularnewline
 Unweighted&355&4&284&-35&608 \tabularnewline
+\textbf{GDP-weighted}&\textbf{350}&\textbf{6}&\textbf{230}&\textbf{-28}&\textbf{558} \tabularnewline
 \midrule &\multicolumn{5}{c}{JMN by country, unweighted} \tabularnewline
@@
 Unweighted average&347&7&254&-16&592 \tabularnewline
-GDP-weighted average&347&2&203&-14&538 \tabularnewline
+\textbf{GDP-weighted average}&\textbf{347}&\textbf{2}&\textbf{203}&\textbf{-14}&\textbf{538} \tabularnewline
 \bottomrule \addlinespace[1.5ex]
@@
 countries.} 
-growth in output is the main driving force behind both waves of growth in
-international trade, contributing to about 56\% (203/365) in
-the first wave and 66\% (361/545) in the second one. It is no coincidence
-if these results are very similar to the ones in table \ref{OR}: 60\% (71/107) in the first globalization and 74\% (183/246) in the second one.
-Insofar as a measure of
-trade costs is defined residually as everything that explains
-differences between domestic and international trade flows, it
-is bound to reword the information contained in an openness
-ratio.
+growth in output is the main driving force behind both waves of growth in international trade, contributing to about 60\% (202/333) in the first wave and 66\% (361/545) in the second one.
+Contrast this with Jack  \textit{et al.}'s results for this sample, respectively 46\% and 63\%.
+It is no coincidence if, despite the sample change, these results are similar to the ones in table \ref{OR}: 62\% (113/182) in the first globalization and 74\% (183/246) in the second one.
+
+Insofar as a measure of trade costs is defined residually as everything that explains differences between domestic and international trade flows, it is bound to reword the information contained in an openness ratio.
 

### 2017-09-15 Paper/Manuscript.tex
réglé un doute

@@
 \usepackage{tabularx}
+\usepackage{tablefootnote}
 
@@
 \end{tabularx}
+
 \caption{Decomposition of the growth in international trade
@@
 trading partners for one country, \emph{our methodl}
-refers to the aggregation method we offer.}\label{glob1}
+refers to the aggregation method we offer. \hspace{\linewidth}
+\textbf{Note:} The contribution of output growth is constant for all countries in our method because it is measured as the growth of total world output}\label{glob1}
 \end{table}
@@
 trading partners for one country, \emph{our methodl}
-refers to the aggregation method we offer.}\label{glob2}
+refers to the aggregation method we offer.\hspace{\linewidth}
+\textbf{Note:} The contribution of output growth is constant for all countries in our method because it is measured as the growth of total world output}\label{glob2}
 \end{table}
@@
 It is no coincidence if, despite the sample change, these results are similar to the ones in table \ref{OR}: 62\% (113/182) in the first globalization and 74\% (183/246) in the second one.
-
-Insofar as a measure of trade costs is defined residually as everything that explains differences between domestic and international trade flows, it is bound to reword the information contained in an openness ratio.
 

### 2017-09-15 Paper/Manuscript.tex
Dernière version

@@
 \usepackage{tablefootnote}
+\usepackage{caption}
+\newcommand{\Note}[1]{\caption*{Note: {#1}} }
 
@@
 \caption{Decomposition of the growth of exportations between
-GDP growth and Openness ratio growth, log differences (interpreted as percentages).
-Figures for Norway are not given because the dissolution of the
+GDP growth and Openness ratio growth, log differences (interpreted as percentages).}
+\Note{Figures for Norway are not given because the dissolution of the
 union between Norway and Sweden (1905) makes them
@@
 
-Let us note $\upsilon_{i}$ the corresponding trade cost faced
-by country $i$ with all its trading partners in the data. Its
-expression is given by Jacks \textit{et al.}'s key equation
-(\ref{keyeq}) applied to a three-countries model:
+Let us note $\upsilon_{i}$ the corresponding trade cost faced by country $i$ with all its trading partners in the data.
+Its expression is given by Jacks \textit{et al.}'s key equation (\ref{keyeq}) applied to a two-countries model:
 
@@
 x_{pi}}\right)^{\frac{1}{2(\sigma-1)}}
-=\left(\frac{x_{ii} \sum_{j \neq i, k \neq i } x_{jk}}{\sum_{j \neq i } x_{ij}\sum_{j \neq i} x_{ji}}\right)^{\frac{1}{2(\sigma-1)}}   \label{keyeq3model}
+=\left(\frac{x_{ii} \sum_{j \neq i, k \neq i } x_{jk}}{\sum_{j \neq i } x_{ij}\sum_{j \neq i} x_{ji}}\right)^{\frac{1}{2(\sigma-1)}}   \label{keyeq2model}
 \end{eqnarray}
@@
 = \sum_{j \neq i}  {\frac{y_i}{\sum_{j \neq i} y_i} \left( \frac{x_{ii} x_{jj}}{x_{ij} x_{ji}}\right)^{\frac{1}{2(\sigma-1)}} }
+ \label{jackssection3}
 \end{eqnarray}
@@
 aggregation method. First wave of globalization,
-1870-1913.  \emph{JMN 2011} refers to the averaging over dyads,
+1870-1913. }
+\Note {\emph{JMN 2011} refers to the averaging over dyads,
 \emph{JMN by country} by country refers to the averaging over
 trading partners for one country, \emph{our methodl}
-refers to the aggregation method we offer. \hspace{\linewidth}
-\textbf{Note:} The contribution of output growth is constant for all countries in our method because it is measured as the growth of total world output}\label{glob1}
+refers to the aggregation method we offer.
+
+The contribution of output growth is constant for all countries in our method because it is measured as the growth of total world output}
+\label{glob1}
 \end{table}
@@
 (logarithms) with \textit{ad hoc} averages and a microfounded
-aggregation method. Second wave of globalization, 1950-2000.
-\emph{JMN 2011} refers to the averaging over dyads,
+aggregation method. Second wave of globalization, 1950-2000.}
+\Note {\emph{JMN 2011} refers to the averaging over dyads,
 \emph{JMN by country} by country refers to the averaging over
 trading partners for one country, \emph{our methodl}
-refers to the aggregation method we offer.\hspace{\linewidth}
-\textbf{Note:} The contribution of output growth is constant for all countries in our method because it is measured as the growth of total world output}\label{glob2}
+refers to the aggregation method we offer.
+
+The contribution of output growth is constant for all countries in our method because it is measured as the growth of total world output}
+\label{glob2}
 \end{table}
@@
 
-Our explanation of why Jacks \textit{et al.}'s aggregate measure of
-trade costs yields different conclusion 
-from table \ref{OR} is because it is sensitive to structure
-effects. To explain this idea, and for clarity purposes, let us consider
-a word where partners are symmetric (domestic trade is equal in $i$ and $j$ and $x_ij$=$x_ji$). From equation \ref{keyeq}, we have:
+The reason why Jacks \textit{et al.}'s aggregate measure of trade costs yields different conclusion the micro-founded method (and openness rate) is its sensitivity to structure
+effects.
+To explain this idea, and for clarity purposes, let us move to a world of symmetric partners (domestic trade is equal in $i$ and $j$ and $x_{ij}$=$x_{ji}$).
+Dyadic bilateral trade flows are not equal because some dyads are further appart than others.
+Equation (\ref{keyeq}) gives:
 
@@
 
-This relation does not of course correspond to the arbitrage
-condition of country $i$, which would involve supply prices,
-but our aim here is to highlight the mechanism behind Jacks et
-al.'s aggregation. Note $\overline{a_j}$ the arithmetic mean
+Note $\overline{a_j}$ the arithmetic mean
 of $a_j$ over $j$ ($\overline{a_j}=1/n \sum_{j=1}^{n}a_j$). One can
@@
 (\ref{tausimplified}) and the measure derived from the
-three-countries model in equation (\ref{keyeq3model}):
+two-countries model in equation (\ref{keyeq2model}):
 
@@
 
-Except for the factor $n$ (which is irrelevant since we are
-concerned with the evolution of the trade costs index) the two
-expressions (\ref{tau}) and (\ref{upsilon}) differ only by the
-mean they use. On the one hand, $\upsilon_i$ uses the
-arithmetic mean of trade flows $x_{ji}$. On the other hand,
-$\tau_i$, because it uses an arithmetic mean over $\tau_{ij}$,
-uses a mean power $1/(1-\sigma)$ of the value of trade flows,
-$\left(\overline{x_{ji}^{\frac{1}{1-\sigma}}}\right)^{1-\sigma}$.
-Appendix \ref{A2} establishes that the curvature
-properties of the function $x \rightarrow
-x^{\frac{1}{1-\sigma}}$, $\sigma>1$ tend to draw the growth of
-$\tau_i$ towards the values of the growth of bilateral trade
-costs incurred with small trading partners.\footnote{The `size' of a country is
-used here to refer to the importance of the volume of its trade
-with $i$ among trade partners of country $i$, regardless of its
-population and GDP.} On the contrary,
-$\upsilon_i$ puts more weight on big trade partners, consistent
-with intuition and their higher share of trade flows.
+Except for the factor $n$ (which is irrelevant since we are concerned with the evolution of the trade costs index) the two equations (\ref{tau}) and (\ref{upsilon}) differ only by the mean they use.
+Equation (\ref{upsilon}) uses the arithmetic mean of trade flows $x_{ji}$.
+Equation (\ref{tau}), because it uses an arithmetic mean over $\tau_{ij}$, uses a power mean of exponent  $1/(1-\sigma)$ of the value of trade flows, $\left(\overline{x_{ji}^{\frac{1}{1-\sigma}}}\right)^{1-\sigma}$.
+Appendix \ref{A2} establishes that the curvature properties of the function $x \rightarrow x^{\frac{1}{1-\sigma}}$, $\sigma>1$ tend to draw the growth of $\tau_i$ towards the values incurred between dyads that do not trade much together (small $x_{ij})$.
+On the contrary, $\upsilon_i$ puts more weight on big trade flows, in a way consistent with intuition.
 %If the
@@
 
-The authors partly correct this biais in the average of $\tau_{ij}$ by the use a end-of-period-GDP
-weighted average. However, as tables \ref{glob1} and \ref{glob2} show,
-weighting by the GDP of trade partners, besides not being
-theoretically justified, does not provide an accurate
-correction of the bias in the measure. One reason for this
-failure is that the relationship between GDP and the importance
-of trade flows is not systematic.\\
+Let us now move away from our symmetric world. Jacks  \textit{et al.} partially correct that bias by using a end-of-period-GDP weights average to compute the average of  $\tau_{ij}$.
+However, as tables \ref{glob1} and \ref{glob2} show, weighting by the GDP of trade partners, besides not being theoretically justified, does not provide an accurate correction of the bias in the measure.
+One reason for this failure is that the relationship between GDP and the importance of trade flows is not systematic.
 
-The average of the bilateral trade costs $\tau_{ij}$ captures both trade
-costs and the distribution of trade costs relative to the size
-of the trading partners.
-Through time, if dyads with small starting trade experience  faster growth of trade than others, the decline of trade costs measured by $\tau_i$ will be overestimated compared to the decline of $\upsilon_i$.
-This fits the data. For both globalizations, figure
-\ref{structure} plots the growth of trade flows (measured by
-$\Delta ln (\sqrt{x_{ij} x_{ji}})$) as a function of the
-initial value of trade (measured by the logarithm of the
-geometric average of bilateral average $ln(\sqrt{x_{ij}
-x_{ji}})$) for all dyads in the sample. This is also confirmed in the contrast between the rate of growth of exports
-(table \ref{OR}) and the rate of growth of bilateral trade flows (table \ref{glob1} and \ref{glob2}), as export growth was relatively faster than bilateral trade growth in the first globalization compared to the second (486/178 > 473/246).
-It confirms that the pairs of countries that initially traded little together experienced faster growth of trade compared to other pairs in the first than in the second globalization.
-This in an interesting result, but not the one presented by JMN.
+For these reasons, Jacks  \textit{et al.}'s average of the bilateral trade costs $\tau_{ij}$ captures both the evolution of trade costs and the distribution of the evolution of trade costs over dyad sizes.
+As a result, if dyads with small starting trade experience faster growth of trade than others, the decline of trade costs measured by $\tau_i$ will be overestimated compared to the decline of $\upsilon_i$.
+This fits the data and can explain the difference they find between the first and the second globalization.
+In the first globalization, pairs of countries that initially traded little together experienced relative faster trade growth.
+For both globalizations, figure \ref{structure} plots the growth of trade flows (measured by $\Delta ln (\sqrt{x_{ij} x_{ji}})$) as a function of the initial value of trade (measured by the logarithm of the geometric average of bilateral average $ln(\sqrt{x_{ij} x_{ji}})$) for all dyads in the sample.
+
+To sum up : in the first globalization, pairs of countries that initially traded little together experienced relative faster trade growth.
+This leads Jacks  \textit{et al.} to overestimate the role of trade costs in the first globalization.
 
@@
 \includegraphics{structure.eps}
-\caption{Growth of trade flows (logarithms) depending on their initial value for the first (1870-1913) and second (1950-2000) globalizations (logarithms).}\label{structure}
+\caption{Growth of trade flows (logarithms) depending on their initial value for the first (1870-1913) and second (1950-2000) globalizations (logarithms).}
+\Note {The correlation for 1870-1913 is -0.635, with 95\% confidence interval: [-0.728;-0.520]
+
+The correlation for 1950-2000 is -0.312, with 95\% confidence interval: [-0.459;-0.147]
+
+These correlation coefficients are statistically significantly different.}
+\label{structure}
 \end{figure}
@@
 trade costs loses part of its interest if there are no causal
-alternatives to explain trade patterns. Such a definition is
+alternatives to explain changing trade patterns. Such a definition is
 therefore bound to reword the information given by trade

### 2017-09-26 Paper/Manuscript.tex
Update Manuscript.tex

@@
 They argue that the reduction in trade costs was the main driving force of trade growth during the first globalization (1870-1913), whereas economic expansion was the main driving force during the second globalization (1950-2000).
-This potentially important result is driven by the use of an \textit{ad hoc} aggregation method of bilateral trade costs to the country and global level.
+This potentially major result is driven by the use of an \textit{ad hoc} aggregation method of bilateral trade costs to the country and global level.
 What \cite{JMN2011} capture is the difference in the relative starting trade of dyads experiencing faster trade growth in the first and second globalization.

### 2017-09-26 Paper/Manuscript.tex
Update Manuscript.tex

@@
 What \cite{JMN2011} capture is the difference in the relative starting trade of dyads experiencing faster trade growth in the first and second globalization.
-More generally, we cast doubts on the possibility to reach conclusions on aggregate costs with a method that excludes a priori changes in non-trade costs determinants of openness rates and hence can only rephrases the information contained in them.
+More generally, we cast doubts on the possibility to reach conclusions on aggregate costs with a method that excludes \textit{a priori} changes in non-trade costs determinants of openness rates and hence can only rephrase the information contained in them.
 \end{abstract}
@@
 This is potentially an important result that sheds light on the globalization processes.
-However, that result is actually driven by an \textit{ad hoc} method of aggregation that captures structure effects.
-The authors use an weighted arithmetic average of bilateral transport costs to compute country-specific trade costs.
-Putting aside the weight, in a simple symmetric three-country model, we show that this is the equivalent to using a power or generalized mean with exponent $1/(1-\sigma)$ on the value of dyadic trade flows ($\sigma > 1$ is elasticity of substitution).
+However, this result is actually driven by an \textit{ad hoc} method of aggregation that captures structure effects.
+The authors use a weighted arithmetic average of bilateral transportation costs to compute country-specific trade costs.
+Putting aside the weight, in a simple symmetric three-country model, we show that this is the equivalent to using a power or generalized mean with exponent $1/(1-\sigma)$ on the value of dyadic trade flows\footnote{namely trade flows between a pair of countries} ($\sigma > 1$ is elasticity of substitution).
 In contrast, a theoretically justified method would use a simple arithmetic mean of the value of the dyadic trade flow. 
@@
 This is not compensated by the weight they use (end-of-period GDP).
-Their conclusion on the difference between the two globalization comes from the fact that the dyads with the fastest growing trade in the first wave of globalization start with very small trade and this this is not the case in the second wave of globalization.
+Their conclusion on the difference between the two globalization periods comes from the fact that the dyads with the fastest growing trade in the first wave of globalization start with very small trade and this this is not the case in the second wave of globalization.
 

### 2017-09-26 Paper/Manuscript.tex
Update Manuscript.tex

@@
 This is not compensated by the weight they use (end-of-period GDP).
-Their conclusion on the difference between the two globalization periods comes from the fact that the dyads with the fastest growing trade in the first wave of globalization start with very small trade and this this is not the case in the second wave of globalization.
+Their conclusion on the difference between the two globalization periods comes from the fact that the dyads with the fastest growing trade in the first wave of globalization start with very small trade; this is not the case in the second wave of globalization.
 
-In this paper, we use a three-countries version of \cite{AW2003}'s model for the aggregation of bilateral trade costs shows that there was no difference in nature between the two globalizations to explain the issue.
-More generally, we cast doubt on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies only on the study of trade flows and excludes a priori other possibles causes for the evolution of openness rates, like the evolution of vertical specialization and changes in the elasticity of substitution between domestic and foreign goods.
+In this paper, we use a three-country version of \cite{AW2003}'s model for the aggregation of bilateral trade costs and show that there was no difference in nature between the two globalizations.
+More generally, we cast doubts on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies solely on the study of trade flows and excludes \textit{a priori} other possible causes for the evolution of openness rates, like the evolution of vertical specialization and changes in the elasticity of substitution between domestic and foreign goods.
 Deducing trade costs from trade flows, and then using trade costs to explain trade flows is essentially a circular reasoning.

### 2017-09-26 Paper/Manuscript.tex
Update Manuscript.tex

@@
 In this paper, we use a three-country version of \cite{AW2003}'s model for the aggregation of bilateral trade costs and show that there was no difference in nature between the two globalizations.
-More generally, we cast doubts on the possibility to distinguish the effect of trade costs and economic expansion through an approach that relies solely on the study of trade flows and excludes \textit{a priori} other possible causes for the evolution of openness rates, like the evolution of vertical specialization and changes in the elasticity of substitution between domestic and foreign goods.
+More generally, we cast doubts on the possibility to distinguish between the impact of trade costs and the impact of economic expansion through an approach that relies solely on the study of trade flows and excludes \textit{a priori} other possible causes for the evolution of openness rates, like the evolution of vertical specialization and changes in the elasticity of substitution between domestic and foreign goods.
 Deducing trade costs from trade flows, and then using trade costs to explain trade flows is essentially a circular reasoning.

### 2017-09-26 Paper/Manuscript.tex
Un peu plus sur x22

--- 2017-09-26 Paper/Manuscript.tex
@@
 The authors thank Marc Adam, Marcelo Olarreaga and anonymous referees for their comments and questions.
-Marc Adam pointed to us a slight data incoherence in Jacks \textit{et al.}'s .dta files that we have corrected in the following analysis.}}
+Marc Adam pointed to us a slight data incoherence in Jacks \textit{et al.}'s .dta files that we have corrected in the following analysis (and helped us with the programming).
+The authors are the sole responsible of mistakes in the text.
+Stata and tex files are available at https://github.com/gdaudin/GT\_ENSAE\_JMN.}}
 \affil[1] {ENSAE, French National School of Statistics and Economic Administration, 3, avenue Pierre Larousse, 92245 Malakoff, FRANCE}
@@
 They argue that the reduction in trade costs was the main driving force of trade growth during the first globalization (1870-1913), whereas economic expansion was the main driving force during the second globalization (1950-2000).
-This potentially major result is driven by the use of an \textit{ad hoc} aggregation method of bilateral trade costs to the country and global level.
+This potentially important result is driven by the use of an \textit{ad hoc} aggregation method of bilateral trade costs to the country and global level.
 What \cite{JMN2011} capture is the difference in the relative starting trade of dyads experiencing faster trade growth in the first and second globalization.
-More generally, we cast doubts on the possibility to reach conclusions on aggregate costs with a method that excludes \textit{a priori} changes in non-trade costs determinants of openness rates and hence can only rephrase the information contained in them.
+More generally, we cast doubts on the possibility to reach conclusions on aggregate costs with a method that excludes a priori changes in non-trade costs determinants of openness rates and hence can only rephrases the information contained in them.
 \end{abstract}
@@
 This is potentially an important result that sheds light on the globalization processes.
-However, this result is actually driven by an \textit{ad hoc} method of aggregation that captures structure effects.
-The authors use a weighted arithmetic average of bilateral transportation costs to compute country-specific trade costs.
-Putting aside the weight, in a simple symmetric three-country model, we show that this is the equivalent to using a power or generalized mean with exponent $1/(1-\sigma)$ on the value of dyadic trade flows\footnote{namely trade flows between a pair of countries} ($\sigma > 1$ is elasticity of substitution).
+However, that result is actually driven by an \textit{ad hoc} method of aggregation that captures structure effects.
+The authors use an weighted arithmetic average of bilateral transport costs to compute country-specific trade costs.
+Putting aside the weight, in a simple symmetric three-country model, we show that this is the equivalent to using a power or generalized mean with exponent $1/(1-\sigma)$ on the value of dyadic trade flows ($\sigma > 1$ is elasticity of substitution).
 In contrast, a theoretically justified method would use a simple arithmetic mean of the value of the dyadic trade flow. 
@@
 This is not compensated by the weight they use (end-of-period GDP).
-Their conclusion on the difference between the two globalization periods comes from the fact that the dyads with the fastest growing trade in the first wave of globalization start with very small trade and this this is not the case in the second wave of globalization.
+Their conclusion on the difference between the two globalization comes from the fact that the dyads with the fastest growing trade in the first wave of globalization start with very small trade and this this is not the case in the second wave of globalization.
 
@@
 \midrule &\multicolumn{5}{c}{Our method} \tabularnewline
-FRA&347&-1&317&-16&648 \tabularnewline
-UK&347&-46&236&-15&522 \tabularnewline
-USA&347&0&192&-13&526 \tabularnewline
-Unweighted average&347&7&254&-16&592 \tabularnewline
-\textbf{GDP-weighted average}&\textbf{347}&\textbf{2}&\textbf{203}&\textbf{-14}&\textbf{538} \tabularnewline
+FRA&347&-1&318&-16&648 \tabularnewline
+UK&347&-46&237&-15&522 \tabularnewline
+USA&347&0&190&-11&526 \tabularnewline
+Unweighted average&347&7&254&-15&592 \tabularnewline
+\textbf{GDP-weighted average}&\textbf{347}&\textbf{2}&\textbf{202}&\textbf{-13}&\textbf{538} \tabularnewline
 \bottomrule \addlinespace[1.5ex]

### 2017-09-26 Paper/Manuscript.tex
Update Manuscript.tex

--- 2017-09-26 Paper/Manuscript.tex
@@
 
-This paper starts by presenting Jacks \textit{et al.}'s approach of the
+In this paper, we first present Jacks \textit{et al.}'s approach of the
 measure of trade costs, and insists on its relevance at the
 bilateral level to control for multilateral trade barriers in
-gravity regressions (section \ref{1}). It then highlights that the
-result on a difference of nature between the two globalizations
+gravity regressions (section \ref{1}). We then highlight that the
+result on the difference of nature between the two globalizations
 is paradoxical since it cannot be deduced from a comparison of the
@@
 
-Where $s_i=y_i/(y_i+y_j)$.
+where $s_i=y_i/(y_i+y_j)$.\\
 
@@
 
-Examining this equation, it is clear that the method consists in determining the evolution of one unobservable variable (trade costs) based on the evolution of two groups of observable variables (trade flows and output) and a number of hypothesis on preferences and parameters ($\sigma$ and the relationship between $x_{ii}$, output and trade).
+Examining this equation, it is clear that the method consists in determining the evolution of one unobservable variable (trade costs) based on the evolution of two groups of observable variables (trade flows and output) and a number of hypotheses on preferences and parameters ($\sigma$ and the relationship between $x_{ii}$, output and trade).
 If one assumes that preferences and parameters do not change (though these points are partly addressed  in appendix B of  \cite{JMN2011}), all changes in trade costs must restate information contained in changes in the measure of trade flows relative to economic size, like an openness ratio defined for instance as the ratio of exportations to GDP.
@@
 \emph{JMN by country} by country refers to the averaging over
-trading partners for one country, \emph{our methodl}
+trading partners for one country, \emph{our method}
 refers to the aggregation method we offer.
@@
 
-The reason why Jacks \textit{et al.}'s aggregate measure of trade costs yields different conclusion the micro-founded method (and openness rate) is its sensitivity to structure
+The reason why Jacks \textit{et al.}'s aggregate measure of trade costs yields different conclusions compared to the micro-founded method (and openness rate) is its sensitivity to structure
 effects.
 To explain this idea, and for clarity purposes, let us move to a world of symmetric partners (domestic trade is equal in $i$ and $j$ and $x_{ij}$=$x_{ji}$).
-Dyadic bilateral trade flows are not equal because some dyads are further appart than others.
+Dyadic bilateral trade flows are not equal because some dyads are further apart than others.
 Equation (\ref{keyeq}) gives:
@@
 (\ref{tausimplified}) and the measure derived from the
-two-countries model in equation (\ref{keyeq2model}):
+two-country model in equation (\ref{keyeq2model}):
 
@@
 
-Let us now move away from our symmetric world. Jacks  \textit{et al.} partially correct that bias by using a end-of-period-GDP weights average to compute the average of  $\tau_{ij}$.
+Let us now move away from our symmetric world. Jacks  \textit{et al.} partially correct that bias by using a end-of-period-GDP weighted average to compute the average of  $\tau_{ij}$.
 However, as tables \ref{glob1} and \ref{glob2} show, weighting by the GDP of trade partners, besides not being theoretically justified, does not provide an accurate correction of the bias in the measure.
@@
 This fits the data and can explain the difference they find between the first and the second globalization.
-In the first globalization, pairs of countries that initially traded little together experienced relative faster trade growth.
-For both globalizations, figure \ref{structure} plots the growth of trade flows (measured by $\Delta ln (\sqrt{x_{ij} x_{ji}})$) as a function of the initial value of trade (measured by the logarithm of the geometric average of bilateral average $ln(\sqrt{x_{ij} x_{ji}})$) for all dyads in the sample.
+For both globalizations, figure \ref{structure} plots the growth of trade flows (measured by $\Delta ln (\sqrt{x_{ij} x_{ji}})$) as a function of the initial value of trade (measured by the logarithm of the geometric average of bilateral average $ln(\sqrt{x_{ij} x_{ji}})$) for all dyads in the sample.\\
 

### 2017-09-26 Paper/Manuscript.tex
Update Manuscript.tex

@@
 They argue that the reduction in trade costs was the main driving force of trade growth during the first globalization (1870-1913), whereas economic expansion was the main driving force during the second globalization (1950-2000).
-This potentially major result is driven by the use of an \textit{ad hoc} aggregation method of bilateral trade costs to the country and global level.
+This potentially major result is driven by the use of an \textit{ad hoc} aggregation method of bilateral trade costs at the country and at the global levels.
 What \cite{JMN2011} capture is the difference in the relative starting trade of dyads experiencing faster trade growth in the first and second globalization.
@@
 openness ratio when it is used to calculate aggregate trade
-costs. This is because it only relates the two through an
-equality between MRS and price ratio. It appears more clearly
+costs. Namely it only relates the two through an
+equality between the MRS and the price ratio. It appears more clearly
 when comparing the \textit{ad hoc} aggregation of bilateral trade costs
@@
 costs used by the authors. In such a model, costs are anything
-that causes consumptions from different countries' products not
-to be equal. To be sure, it is essential to highlight that
+that causes consumption flows from different countries' products not
+to be equal. Yet it is essential to highlight that
 trade costs cannot be reduced to tariffs or transportation

### 2017-09-26 Paper/Manuscript.tex
Update Manuscript.tex

@@
 This potentially major result is driven by the use of an \textit{ad hoc} aggregation method of bilateral trade costs at the country and at the global levels.
-What \cite{JMN2011} capture is the difference in the relative starting trade of dyads experiencing faster trade growth in the first and second globalization.
+What \cite{JMN2011} capture is that some pair of countries (dyads) experienced faster trade growth in the first globalization than in the the second globalization.
 More generally, we cast doubts on the possibility to reach conclusions on aggregate costs with a method that excludes \textit{a priori} changes in non-trade costs determinants of openness rates and hence can only rephrase the information contained in them.
@@
 It is surprising because it is not clear what is added by their method to the examination of the openness rates.
-As the next section shows, this contrast it is actually driven by the \textit{ad hoc} way they aggregate trade costs.
+As the next section shows, this contrast is actually driven by the \textit{ad hoc} way they aggregate trade costs.
 

### 2017-09-28 Paper/Manuscript.tex
Add files via upload

@@
 However, this result is actually driven by an \textit{ad hoc} method of aggregation that captures structure effects.
-The authors use a weighted arithmetic average of bilateral transportation costs to compute country-specific trade costs.
-Putting aside the weight, in a simple symmetric three-country model, we show that this is the equivalent to using a power or generalized mean with exponent $1/(1-\sigma)$ on the value of dyadic trade flows\footnote{namely trade flows between a pair of countries} ($\sigma > 1$ is elasticity of substitution).
-In contrast, a theoretically justified method would use a simple arithmetic mean of the value of the dyadic trade flow. 
+The authors use a weighted arithmetic average of their measure of bilateral trade costs to compute country-specific trade costs.
+Putting aside the weight, this is equivalent to using a power mean with exponent $1/(1-\sigma)$ of the values of dyadic trade flows ($\sigma > 1$ is the elasticity of substitution).
+In contrast, we show that an aggregation method theoretically rooted in \cite{AW2003}'s model would use a simple arithmetic mean of the value of the dyadic trade flows. 
 Because $1/(1-\sigma)$ < 0, the importance of small dyadic trade flows in the computation of country-specific trade costs is too large in the authors' computations.
 This is not compensated by the weight they use (end-of-period GDP).
-Their conclusion on the difference between the two globalization periods comes from the fact that the dyads with the fastest growing trade in the first wave of globalization start with very small trade; this is not the case in the second wave of globalization.
+The authors' conclusion on the difference between the two globalization periods comes from the fact that the dyads with the fastest growing trade in the first wave of globalization start with very small trade; this is not the case in the second wave of globalization.
+Indeed, we show that using our theory-based aggregation method, there is no difference in nature between the two globalizations.
 
-In this paper, we use a three-country version of \cite{AW2003}'s model for the aggregation of bilateral trade costs and show that there was no difference in nature between the two globalizations.
 More generally, we cast doubts on the possibility to distinguish between the impact of trade costs and the impact of economic expansion through an approach that relies solely on the study of trade flows and excludes \textit{a priori} other possible causes for the evolution of openness rates, like the evolution of vertical specialization and changes in the elasticity of substitution between domestic and foreign goods.
@@
 
-In this paper, we first present Jacks \textit{et al.}'s approach of the
+We first present Jacks \textit{et al.}'s approach to the
 measure of trade costs, and insists on its relevance at the
 bilateral level to control for multilateral trade barriers in
-gravity regressions (section \ref{1}). We then highlight that the
+gravity regressions. We then highlight that the
 result on the difference of nature between the two globalizations
 is paradoxical since it cannot be deduced from a comparison of the
-evolution of openness ratios. Section \ref{2} shows that the
+evolution of openness ratios (section \ref{1}). Section \ref{2} shows that the
 conclusion is only driven by the authors' \textit{ad hoc} aggregation
@@
 bilateral trade does not depend on bilateral trade barriers \textit{per
-se}, but bilateral trade barriers relative to impediments to
-trade with all other trading partners. \cite{AW2003} show that the
+se}, but bilateral trade barriers relative to trade barriers with all other trading partners. \cite{AW2003} show that the
 equilibrium imposes the following relation:
@@
 
-When departing from \cite{AW2003}'s multilateral resistances, the arbitrage condition Jacks \textit{et al.} base their computation on is an equality between a Marginal Rate of Substitution (MRS) and a price ratio for a CES utility function, or more precisely the product of two such equalities, one from the program of each trade partner.
+When departing in this way from \cite{AW2003}'s multilateral resistances, the arbitrage condition Jacks \textit{et al.} base their computation on is an equality between a Marginal Rate of Substitution (MRS) and a price ratio for a CES utility function---or more precisely the product of two such equalities, one from the program of each trade partner.
 Appendix \ref{A1} derives equation (\ref{keyeq}) this way without passing through multilateral resistance.
@@
 However, Jacks, Meissner and Novy also suggest in their article that \cite{AW2003}'s model can be used to move away from the gravity regression approach.
-Instead of explaining trade flows by observable proxies for costs, they quantify the impact of all impediments to trade:
+Instead of explaining trade flows by observable proxies for costs, they quantify the impact of all impediments to trade from trade flows:
 \emph{``We, therefore, infer trade costs from trade flows.
@@
 
-Examining this equation, it is clear that the method consists in determining the evolution of one unobservable variable (trade costs) based on the evolution of two groups of observable variables (trade flows and output) and a number of hypotheses on preferences and parameters ($\sigma$ and the relationship between $x_{ii}$, output and trade).
+The method consists in determining the evolution of one unobservable variable (trade costs) based on the evolution of two groups of observable variables (trade flows and output) and a number of hypotheses on preferences and parameters ($\sigma$ and the relationship between $x_{ii}$, output and trade).
+A method that uses data on trade flows and output alone can only teach us so much about their determinants since non-trade costs determinants of trade flows (e.g.
+the elasticity of substitution and the importance of vertical specialization) are treated as parameters.
 If one assumes that preferences and parameters do not change (though these points are partly addressed  in appendix B of  \cite{JMN2011}), all changes in trade costs must restate information contained in changes in the measure of trade flows relative to economic size, like an openness ratio defined for instance as the ratio of exportations to GDP.
 
-A method that uses data on trade flows and output alone can only teach us so much about their determinants as non-trade costs determinants of trade flows (e.g.
-the elasticity of substitution and the importance of vertical specialization) are treated as parameters.
-
-Still, when we decompose the level of exportations of a country simply as the product of its GDP and its openness ratio (see table \ref{OR}) we find that the increase in exportations is mainly explained by GDP growth for the large majority of the 27 countries in the data.\footnote{We follow the authors in interpreting log differences as percentages.
+Yet, Jacks \emph{et al.}'s results seem to show otherwise.
+When we decompose the level of exportations of a country simply as the product of its GDP and its openness ratio (see table \ref{OR}) we find that for both globalizations the increase in exportations is mainly explained by GDP growth for the large majority of the 27 countries in the data.\footnote{We follow the authors in interpreting log differences as percentages.
 However, one should keep in mind that given the size of the changes, this is a very inexact approximation.}
@@
 It is surprising because it is not clear what is added by their method to the examination of the openness rates.
-As the next section shows, this contrast is actually driven by the \textit{ad hoc} way they aggregate trade costs.
-
+As the next section shows, this difference is actually driven by the \textit{ad hoc} way they aggregate the measure of bilateral trade costs.
 \begin{table}[tbp] \centering
@@
 
-\cite{AW2003}'s model can be used to calculate the aggregate
-trade costs a single country faces in its trade with all
-its trading partners. Indeed, using a two-countries model where
-all trade partners of country $i$ are treated as one single
-country provides an aggregate measure of trade costs that does
-not rely on any \textit{ad hoc} aggregation method but instead on
-microfoundations. The drawback of this method is that it is very data-intensive. It requires the knowledge of all bilateral trade flows between the countries in the data.
+To show this point, we reproduce Jacks \emph{et al.}'s decomposition using instead an aggregation method theoretically derived from \cite{AW2003}'s model.
+We calculate the aggregate trade costs a single country faces with all its trading partners by considering the two-country version of the model where all trade partners of country $i$ are treated as a unique one.
+
+%The last sentence ""The drawback of..." seemed peripheral and a distraction from the main line of thought? Besides, it is not seem more data-intensive than Jacks et al's measure
 
@@
 now
-includes cross-border trade. \\
-
-Compare this with Jacks \textit{et al.}'s  method to measure the mean trade cost faced by a country : 
+includes cross-border trade. Compare this with Jacks \textit{et al.}'s  method to measure the mean trade cost faced by a country : 
 \begin{eqnarray}
@@
 
-To illustrate the practical difference between these two methods, we provide the same decomposition in four terms of the increase in trade flows during the two globalizations as in the authors' article for the countries with the most trade partners available in the data (France (24), the UK (25), the USA (23)).
+Tables \ref{glob1} and \ref{glob2} provide the decomposition in four terms of the increase in trade flows during the two globalizations for our and Jacks \textit{et al.}'s aggregation methods, for the countries with the most trade partners available in the data (France (24), the UK (25), the USA (23)).
 This requires studying only trade between Belgium, Canada, France, Germany, India, Italy, Spain, United Kingdom and United States of America, as this is the largest set in the data where all bilateral trade flows are available.
-We also provide the unweighted and end-of-period-GDP-weighted averages for all countries in the sample. 
-Results are displayed in tables \ref{glob1} and \ref{glob2} along with the results with the
-Jacks \textit{et al.}'s method.\footnote{They include unweighted and
+The tables also provide the unweighted and end-of-period-GDP-weighted averages for all countries in the sample.\footnote{The tables include unweighted and
 	end-of-period-GPD-weighted averages over country dyads of
@@
 
-Tables \ref{glob1} and \ref{glob2} highlight how much the decomposition
-between the decrease in trade costs and the income growth depends on the aggregation technique.
-When we use the microfounded aggregation method,\footnote{We selected the (GDP-weighted) average of our measure in order
+Comparing the results for the two methods highlights how much the decomposition
+between the decrease in trade costs and income growth depends on the aggregation method.
+When we use the microfounded aggregation method, growth in output is the main driving force behind both waves of growth in international trade, contributing to about 60\% (202/333) in the first wave and 66\% (361/545) in the second one.\footnote{We selected the (GDP-weighted) average of our measure in order
 	to allow a clear comparison with Jacks \textit{et al.}'s results. There
@@
 	conclusion of this exercice does not depend on averaging over
-countries.} 
-growth in output is the main driving force behind both waves of growth in international trade, contributing to about 60\% (202/333) in the first wave and 66\% (361/545) in the second one.
-Contrast this with Jack  \textit{et al.}'s results for this sample, respectively 46\% and 63\%.
-It is no coincidence if, despite the sample change, these results are similar to the ones in table \ref{OR}: 62\% (113/182) in the first globalization and 74\% (183/246) in the second one.
+	countries.} These results are similar to the decomposition of trade flows between output and openness ratio in table \ref{OR}---62\% and 74\%---despite the sample change.
+Contrast this with Jack  \textit{et al.}'s results, respectively 46\% and 63\%.
 
@@
 
-The reason why Jacks \textit{et al.}'s aggregate measure of trade costs yields different conclusions compared to the micro-founded method (and openness rate) is its sensitivity to structure
+The reason why Jacks \textit{et al.}'s aggregate measure of trade costs yields different conclusions compared to the micro-founded method (and openness rates) is its sensitivity to structure
 effects.
-To explain this idea, and for clarity purposes, let us move to a world of symmetric partners (domestic trade is equal in $i$ and $j$ and $x_{ij}$=$x_{ji}$).
-Dyadic bilateral trade flows are not equal because some dyads are further apart than others.
+To explain this idea, and for clarity purposes, let us move to a world of symmetric partners: domestic trade is equal in $i$ and $j$, $x_{ii}=x_{jj}$, and imports are equal to exports, $x_{ij}$=$x_{ji}$.
+(Obviously, trade flows $x_{ij}$ can still differ across trading partners---otherwise there would be no aggregation problem).
 Equation (\ref{keyeq}) gives:
+% The last sentence "Dyadic bilateral trade flows are not equal because some dyads are further apart than others." seemed out of context?
 
@@
 of $a_j$ over $j$ ($\overline{a_j}=1/n \sum_{j=1}^{n}a_j$). One can
-then compare the unweighted average of $\tau_{ij}$ according to equation 
+write the unweighted average of $\tau_{ij}$ according to equation 
 (\ref{tausimplified}) and the measure derived from the
-two-country model in equation (\ref{keyeq2model}):
+two-country model in equation (\ref{keyeq2model}) as:
 
@@
 Equation (\ref{tau}), because it uses an arithmetic mean over $\tau_{ij}$, uses a power mean of exponent  $1/(1-\sigma)$ of the value of trade flows, $\left(\overline{x_{ji}^{\frac{1}{1-\sigma}}}\right)^{1-\sigma}$.
-Appendix \ref{A2} establishes that the curvature properties of the function $x \rightarrow x^{\frac{1}{1-\sigma}}$, $\sigma>1$ tend to draw the growth of $\tau_i$ towards the values incurred between dyads that do not trade much together (small $x_{ij})$.
-On the contrary, $\upsilon_i$ puts more weight on big trade flows, in a way consistent with intuition.
+Appendix \ref{A2} establishes that the curvature properties of the function $x \rightarrow x^{\frac{1}{1-\sigma}}$, $\sigma>1$ tend to draw the growth of $\tau_i$ towards the values incurred with the trading partners with which $i$ does not trade much (small $x_{ij})$.
+On the contrary, $\upsilon_i$ puts more weight on the trading partners with big trade flows, simply because they account for a larger part of trade.
 %If the
@@
 equality between the MRS and the price ratio. It appears more clearly
-when comparing the \textit{ad hoc} aggregation of bilateral trade costs
-to an aggregated measure of trade costs based on
-\cite{AW2003}'s model. Still, if their method fails to offer a
+when replacing the \textit{ad hoc} aggregation of bilateral trade costs
+with an aggregation method directly rooted in \cite{AW2003}'s model. Still, if their method fails to offer a
 full alternative to traditional investigations of trade costs,
@@
 and convexity properties of the function $\phi$. For instance,
-if $\phi$ is convex, the convexity inequality gives:
+if $\phi$ is convex, Jensen inequality gives:
 
@@
 If $\phi$, and hence $\phi^{-1}$, is also increasing, the
-$\phi$-mean is superior to the arithmetic mean. Intuitively,
+$\phi$-mean is greater than the arithmetic mean. Intuitively,
 the convexity of $\phi$ gives, relative to the arithmetic mean,
@@
 $\phi$ the power functions $\phi: x \rightarrow
-x^{\frac{1}{1-\rho}}$, $\rho>1$ where the coherence with the
+x^{\frac{1}{1-\sigma}}$, $\sigma>1$ where consistency with the
 model imposes an arithmetic mean. Hence, since $\phi$ is
-decreasing and convex, the substitution tends to overweight
-small countries.\\
+decreasing and convex, the substitution tends to underestimate trade costs.\\
 
-This is a static result. We are interested in the dynamic
-behavior of trade flows. But Jacks \textit{et al.}'s measure also
-overweights initially small trade partners in terms of the
-increase in trade flows. This can be seen by calculating the
-elasticities to the importations from country $k$ ($a_k$) of
+This is a static result. We are interested in its consequences for the dynamic
+behavior of trade flows. In terms of the
+increase in trade flows, Jacks \textit{et al.}'s measure
+overweights initially small trade partners. This can be seen by calculating the
+elasticities to the importations from a country $k$ ($a_k$) of
 the arithmetic mean ($m$) and of Jacks \textit{et al.}'s mean
@@
 Jacks \textit{et al.}'s measure is biased toward the growth rates of
-costs incurred with small trade partners.
+costs incurred with initially small trade partners.
 

### 2017-09-28 Paper/Manuscript.tex
Dernière modifications

@@
 The authors use a weighted arithmetic average of their measure of bilateral trade costs to compute country-specific trade costs.
-Putting aside the weight, this is equivalent to using a power mean with exponent $1/(1-\sigma)$ of the values of dyadic trade flows ($\sigma > 1$ is the elasticity of substitution).
+This is equivalent to using a power mean with exponent $1/(1-\sigma)$ of the values of dyadic trade flows ($\sigma > 1$ is the elasticity of substitution) to aggregate dyadic costs into country-specific costs.
 In contrast, we show that an aggregation method theoretically rooted in \cite{AW2003}'s model would use a simple arithmetic mean of the value of the dyadic trade flows. 
@@
 
-More generally, we cast doubts on the possibility to distinguish between the impact of trade costs and the impact of economic expansion through an approach that relies solely on the study of trade flows and excludes \textit{a priori} other possible causes for the evolution of openness rates, like the evolution of vertical specialization and changes in the elasticity of substitution between domestic and foreign goods.
-Deducing trade costs from trade flows, and then using trade costs to explain trade flows is essentially a circular reasoning.
-Therefore, Jacks \textit{et al.}'s approach cannot be an alternative to traditional investigations of impediments to trade, such as commodity price gaps.
+More generally, we cast doubts on the possibility to distinguish between the impact of aggregate trade costs and the impact of aggregate economic expansion through an approach that relies solely on the study of trade flows and excludes \textit{a priori} other possible causes for the evolution of openness rates, like the evolution of vertical specialization and changes in the elasticity of substitution between domestic and foreign goods.
+Therefore, Jacks \textit{et al.}'s approach cannot be an alternative to traditional investigations of impediments to trade at the global level, such as commodity price gaps.
+It is much more useful to study bilateral trade costs, even though its usefulness is limited by the amount of structure that must be imposed on the data to use it.
 
@@
 difference in the distribution of trade growth over trading
-partners.
+dyads.
 
-\section{\label{1} Deducing trade costs from trade flows, explaining trade flows with trade costs}
+\section{\label{1} Deducing trade costs from trade flows}
 
@@
 country $i$ and can be interpreted as multilateral trade
-barriers, and $t_{ij}$ is the trade costs factor between $i$
+barrier or resistance, and $t_{ij}$ is the trade costs factor between $i$
 and $j$. Trade costs factors are assumed to be symmetric, i.e.
@@
 
-The last equality defines $\tau_{ij}$, the geometric mean of trade costs from country $i$ to $j$ and $j$ to $i$ \emph{relative to intra-national trade costs in countries $i$ and $j$} (no assumption of symmetry of bilateral trade costs is imposed).
+The last equality defines $\tau_{ij}$, trade costs from country $i$ to $j$ and $j$ to $i$ \emph{relative to intra-national trade costs in countries $i$ and $j$} (no assumption of symmetry of bilateral trade costs is imposed).
 It is the trade cost measure used by the authors.
@@
 
-However, Jacks, Meissner and Novy also suggest in their article that \cite{AW2003}'s model can be used to move away from the gravity regression approach.
+Jacks, Meissner and Novy also suggest in their article that \cite{AW2003}'s model can be used to move away from the gravity regression approach.
 Instead of explaining trade flows by observable proxies for costs, they quantify the impact of all impediments to trade from trade flows:
@@
 the elasticity of substitution and the importance of vertical specialization) are treated as parameters.
-If one assumes that preferences and parameters do not change (though these points are partly addressed  in appendix B of  \cite{JMN2011}), all changes in trade costs must restate information contained in changes in the measure of trade flows relative to economic size, like an openness ratio defined for instance as the ratio of exportations to GDP.
+If one assumes that preferences and parameters do not change (though these points are partly addressed  in appendix B of  \cite{JMN2011}), all changes in trade costs must restate information contained in changes in the measure of trade flows relative to economic size.
+Despite this heavy reliance on a priori hypotheses, or structure, it is an useful approach at the bilateral level because there no obvious simple indicator of bilateral openness.
+At the agregate level, there are obvious simple indicators of openness (e.g.
+openness ratio defined as the ratio of exportations to GDP), and the result of \cite{JMN2011} should restate it.
 
 Yet, Jacks \emph{et al.}'s results seem to show otherwise.
-When we decompose the level of exportations of a country simply as the product of its GDP and its openness ratio (see table \ref{OR}) we find that for both globalizations the increase in exportations is mainly explained by GDP growth for the large majority of the 27 countries in the data.\footnote{We follow the authors in interpreting log differences as percentages.
+When one decomposes the level of exportations of a country simply as the product of its GDP and its openness ratio (see table \ref{OR}) one finds that for both globalizations the increase in exportations is mainly explained by GDP growth for the large majority of the 27 countries in the data.\footnote{We follow the authors in interpreting log differences as percentages.
 However, one should keep in mind that given the size of the changes, this is a very inexact approximation.}
@@
 This contrast between both decompositions of the growth of exports is at the center of the argument by Jack \textit{et al.}.
-It is surprising because it is not clear what is added by their method to the examination of the openness rates.
-As the next section shows, this difference is actually driven by the \textit{ad hoc} way they aggregate the measure of bilateral trade costs.
+It is surprising because it is not clear what is added by their method to the examination of openness rates.
+As the next section shows, it is actually driven by the \textit{ad hoc} way they aggregate their measure of bilateral trade costs.
 \begin{table}[tbp] \centering
@@
 measure that provides a summary statistic of the
-evolution of trade costs across all dyads in the sample, as
-well as for different regions of the world. To move from
+evolution of trade costs across all dyads in the sample. To move from
 bilateral costs $\tau_{ij}$ to an aggregate measure of trade
@@
 There is no justification for this aggregation method. We argue that
-all the results they reach that are not a reformulation of the
+all the agregate results they reach that are not a reformulation of the
 evolution of the openness ratio during the two
@@
 
-%The last sentence ""The drawback of..." seemed peripheral and a distraction from the main line of thought? Besides, it is not seem more data-intensive than Jacks et al's measure
+%The drawback of this method is that it is very data-intensive as it requires the knowledge of all bilateral trade flows between the countries in the data, whereas Jacks \emph{et al.}'s method
+
+%The last sentence ""The drawback of..." seemed peripheral and a distraction from the main line of thought? Besides, it is not seem more data-intensive than Jacks et al's measure SD
+%Well, theirs is less data intensive as it does not requires the full population of bilateral trade flows. But I agree it can be treated as a distraction GD
 
@@
 
-The reason why Jacks \textit{et al.}'s aggregate measure of trade costs yields different conclusions compared to the micro-founded method (and openness rates) is its sensitivity to structure
+The reason why Jacks \textit{et al.}'s aggregate measure of trade costs yields different conclusions compared to the microfounded method (and openness rates) is its sensitivity to structure
 effects.
@@
 that causes consumption flows from different countries' products not
-to be equal. Yet it is essential to highlight that
+to be equal. We agree that it is essential to highlight that
 trade costs cannot be reduced to tariffs or transportation

### 2017-09-29 Paper/Manuscript.tex
Intégration des dernières remarques

@@
 This potentially major result is driven by the use of an \textit{ad hoc} aggregation method of bilateral trade costs at the country and at the global levels.
-What \cite{JMN2011} capture is that some pair of countries (dyads) experienced faster trade growth in the first globalization than in the the second globalization.
+What \cite{JMN2011} capture is that some pairs of countries experienced faster trade growth in the first globalization than in the second globalization.
 More generally, we cast doubts on the possibility to reach conclusions on aggregate costs with a method that excludes \textit{a priori} changes in non-trade costs determinants of openness rates and hence can only rephrase the information contained in them.
@@
 (1921-1939) thanks to the impressive set of data they collected
-on trade flows, GDP and exports between 27
+on trade flows and GDP between 27
 countries\footnote{Argentina, Australia, Austria, Belgium,
@@
 However, this result is actually driven by an \textit{ad hoc} method of aggregation that captures structure effects.
-The authors use a weighted arithmetic average of their measure of bilateral trade costs to compute country-specific trade costs.
-This is equivalent to using a power mean with exponent $1/(1-\sigma)$ of the values of dyadic trade flows ($\sigma > 1$ is the elasticity of substitution) to aggregate dyadic costs into country-specific costs.
+The authors use a weighted arithmetic average of their measure of trade costs between country pairs (dyads) to compute country-specific trade costs.
+This is equivalent to computing a power mean with exponent $1/(1-\sigma)$ of the values of dyadic trade flows ($\sigma > 1$ is the elasticity of substitution).
 In contrast, we show that an aggregation method theoretically rooted in \cite{AW2003}'s model would use a simple arithmetic mean of the value of the dyadic trade flows. 
@@
 It is much more useful to study bilateral trade costs, even though its usefulness is limited by the amount of structure that must be imposed on the data to use it. 
+Still, without the added structure imposed on the data, deducing trade costs from trade flows, and then using trade costs to explain trade flows would essentially be a circular reasoning
 
 We first present Jacks \textit{et al.}'s approach to the
-measure of trade costs, and insists on its relevance at the
+measure of trade costs, and insist on its relevance at the
 bilateral level to control for multilateral trade barriers in
@@
 goods over one single period. In this Armington world,
-production is not modelized\footnote{In the working paper
+production is not modelized and each country is initially
+endowed with a differentiated representative good. Trade occurs
+because of consumers' taste for diversity.\footnote{In the working paper
 version of \cite{JMN2010}, the authors provide a version of the
 model with production. The key equation is identical to the one
-of the model without production.} and each country is initially
-endowed with a differentiated representative good. Trade occurs
-because of consumers' taste for diversity. The preferences of
+of the model without production.} The preferences of
 all countries are assumed to be identical and modelized by a
@@
 measured in value-added and exportations as gross value are
-addressed in appendix B of \cite{JMN2011}.}. The equation above
+addressed in appendix B of \cite{JMN2011}.} The equation above
 becomes:
@@
 If one assumes that preferences and parameters do not change (though these points are partly addressed  in appendix B of  \cite{JMN2011}), all changes in trade costs must restate information contained in changes in the measure of trade flows relative to economic size.
-Despite this heavy reliance on a priori hypotheses, or structure, it is an useful approach at the bilateral level because there no obvious simple indicator of bilateral openness.
+Despite this reliance on a priori hypotheses, or structure, it is a useful approach at the bilateral level because there no simple indicator of bilateral openness.
 At the agregate level, there are obvious simple indicators of openness (e.g.
@@
 Let us note $\upsilon_{i}$ the corresponding trade cost faced by country $i$ with all its trading partners in the data.
-Its expression is given by Jacks \textit{et al.}'s key equation (\ref{keyeq}) applied to a two-countries model:
+Its expression is given by Jacks \textit{et al.}'s key equation (\ref{keyeq}) applied to a two-country model:
 
